### 2015-08-28 paper.tex
Initial commit

\documentclass{vldb}
\usepackage{graphicx}
\usepackage{balance}

\begin{document}

% title
\title{A Survey on Ordered Binary Decision Diagrams}

% author(s)
\numberofauthors{1}

\author{
\alignauthor
	Daniel Kocher\\
    \affaddr{University of Salzburg}\\
    \email{Daniel.Kocher@stud.sbg.ac.at}
}

\maketitle

\begin{abstract}
Tbd.
\end{abstract}

\section{Introduction}
Tbd.

% balance columns on last page
\balance

%\bibliographystyle{abbrv}
%\bibliography{paper}

\end{document}

### 2015-09-08 paper.tex
Add citations
Add additional references
Add additional citation
Add additional citations
Fix syntax
Add basic structure and first version of abstract

@@
 \begin{abstract}
-Tbd.
+This paper serves as a survey on \textit{ordered binary Decision diagrams}
+(\textit{OBDDs}). It provides an overview of the representation of OBDDs, the
+operations which can be applied efficiently, some aspects when implementing them
+as well as their limitations and how to probably overcome them by using
+alternative representations. No new results or insights are provided.  [TO BE
+REVISED].
 \end{abstract}
@@
 \section{Introduction}
+\label{sec:introduction}
+
 Tbd.
+
+\section{Representation Basics}
+\label{sec:representation-basics}
+
+Tbd.
+
+\subsection{Reducing and Ordering BDDs}
+\label{subsec:reducing-and-ordering-bdds}
+
+Tbd.
+
+\section{Efficient Operations}
+\label{sec:efficient-operations}
+
+Tbd.
+
+\section{Implementational Aspects}
+\label{sec:implementation-aspects}
+
+Tbd.
+
+\subsection{Dynamic Variable Ordering}
+\label{subsec:dynamic-variable-ordering}
+
+Tbd.
+
+\section{Limitations}
+\label{sec:limitations}
+
+Tbd.
+
+\section{Alternative Representations}
+\label{sec:alternative-representations}
+
+Tbd.
+
+\section{Citations (temporary)}
+
+\noindent
+\par
+LEE59\cite{LEE59}, AKERS78\cite{AKERS78}, BRYANT86\cite{BRYANT86},
+\par
+BRACE90\cite{BRACE90}, BRYANT91\cite{BRYANT91}, BRYANT92\cite{BRYANT92},
+\par
+RUDELL93\cite{RUDELL93}, BRYANT95\cite{BRYANT95}, BOLLIG96\cite{BOLLIG96},
+\par
+ANDERSEN97\cite{ANDERSEN97}, WOELFEL01\cite{WOELFEL01}
 
@@
 
-%\bibliographystyle{abbrv}
-%\bibliography{paper}
+\bibliographystyle{abbrv}
+\bibliography{paper}
 

### 2015-09-08 paper.tex
Add introduction

@@
 as well as their limitations and how to probably overcome them by using
-alternative representations. No new results or insights are provided.  [TO BE
-REVISED].
+alternative representations. No new results or insights are provided.
 \end{abstract}
@@
 
-Tbd.
+In computer-aided design (CAD) as well as other domains like artificial
+intelligence or combinatorics, many problems can be formulated as Boolean
+functions. These Boolean functions then can be represented symbolically using the
+\textit{Binary Decision Diagrams} (\textit{BDDs}) introduced by Lee\cite{LEE59}
+and Akers\cite{AKERS78} in 1959 and 1978, respectively.
+
+In 1986 Bryant\cite{BRYANT86} published a paper of high impact, describing a
+\textit{reduced} and \textit{ordered} class of BDDs, so-called \textit{ROBDDs}
+(\textit{Reduced Ordered} BDDs) but in literature mostly referred to as
+\textit{OBDDs}. OBDDs represent Boolean functions in a \textit{canonical}
+form. This results in a compact form and very efficient tests for some properties,
+e.g. satisfiability or equivalence\cite{BRYANT86}.
+
+Already in the inital paper describing OBDDs, the importance of the variable
+ordering was highlighted. Only if the variable ordering is chosen properly, it
+will result in a small graph (OBDD) and, in turn, in more efficient testing.
+
+Nonetheless representing Boolean function using OBDDs has some drawbacks. Classes
+of Boolean functions exist for which the size of the graph grows exponentially,
+e.g. integer multiplication~\cite{BRYANT86, BRYANT91, WOELFEL01}.
+
+These limitations of OBDDs motivated researchers to introduce other
+representations which e.g. relax the ordering or change the interpretation of
+nodes~\cite{BRYANT95}. Such representations may be beneficial for some classes
+of Boolean functions but mostly suffer from losing the canonical form and hence
+the desirable properties OBDDs are famous for.
+
+This paper serves as survey on symbolic Boolean function manipulation using OBDDs.
+The following section describes the representation of Boolean function through
+BDDs as well as OBDDs. The third section describes operations which can be applied
+to OBDDs efficiently. No source is provided but those can be looked up in the
+corresponding papers easily (references are provided for each algorithm).
+Section~\ref{sec:implementation-aspects} describes aspects one has to take into
+account when implementing an OBDD package, such as what data structures to use
+or how to deal with the problem of variable ordering. Limitations of the OBDD
+representation are described in the subsequent section. The last section
+discusses alternative representations to possibly overcome the limitations of
+the OBDD representation.
 
@@
 \par
-LEE59\cite{LEE59}, AKERS78\cite{AKERS78}, BRYANT86\cite{BRYANT86},
+BRYANT86\cite{BRYANT86},
 \par

### 2015-09-08 paper.tex
Add tbr command
Fix typo
Update introduction

@@
 \usepackage{balance}
+
+\newcommand{\tbr}{\textbf{[TO BE REVISED]}}
 
@@
 alternative representations. No new results or insights are provided.
+\tbr
 \end{abstract}
@@
 This paper serves as survey on symbolic Boolean function manipulation using OBDDs.
-The following section describes the representation of Boolean function through
-BDDs as well as OBDDs. The third section describes operations which can be applied
-to OBDDs efficiently. No source is provided but those can be looked up in the
-corresponding papers easily (references are provided for each algorithm).
+The following section describes the representation of Boolean functions through
+BDDs as well as OBDDs. It also highlights the advantages and disadvantages of the
+OBDD representation. The third section describes operations which can be applied
+to OBDDs efficiently. No source code is provided but those can be looked up in
+the corresponding papers easily (references are provided for each algorithm).
 Section~\ref{sec:implementation-aspects} describes aspects one has to take into
@@
 the OBDD representation.
+\tbr
 

### 2015-09-08 paper.tex
Add tbd macro; Fix typo

@@
 \newcommand{\tbr}{\textbf{[TO BE REVISED]}}
+\newcommand{\tbd}{\textbf{[TO BE DONE]}}
 
@@
 
-Nonetheless representing Boolean function using OBDDs has some drawbacks. Classes
+Nonetheless representing Boolean functions using OBDDs has some drawbacks. Classes
 of Boolean functions exist for which the size of the graph grows exponentially,
@@
 These limitations of OBDDs motivated researchers to introduce other
-representations which e.g. relax the ordering or change the interpretation of
-nodes~\cite{BRYANT95}. Such representations may be beneficial for some classes
-of Boolean functions but mostly suffer from losing the canonical form and hence
-the desirable properties OBDDs are famous for.
+representations which e.g. relax the ordering~\cite{BRYANT95} or change the
+interpretation of nodes~\cite{BRYANT95}~\cite{ANDERSEN97}. Such representations
+may be beneficial for some classes of Boolean functions but mostly suffer from
+losing the canonical form and hence the desirable properties OBDDs are famous for.
 
@@
 
-Tbd.
+\tbd
 
@@
 \par
-BRYANT86\cite{BRYANT86},
+BRACE90\cite{BRACE90}, BRYANT92\cite{BRYANT92},
 \par
-BRACE90\cite{BRACE90}, BRYANT91\cite{BRYANT91}, BRYANT92\cite{BRYANT92},
-\par
-RUDELL93\cite{RUDELL93}, BRYANT95\cite{BRYANT95}, BOLLIG96\cite{BOLLIG96},
-\par
-ANDERSEN97\cite{ANDERSEN97}, WOELFEL01\cite{WOELFEL01}
+RUDELL93\cite{RUDELL93}, BOLLIG96\cite{BOLLIG96}
 

### 2015-09-08 paper.tex
Add paragraph to introduction

@@
 as well as their limitations and how to probably overcome them by using
-alternative representations. No new results or insights are provided.
-\tbr
+alternative representations. No new results or insights are provided. \tbr
 \end{abstract}
@@
 
+Bryant also introduced and experimentally evaluated efficient implementation
+techniques for OBDDs in collaboration with Brace and
+Rudell~\cite{BRACE90, BRYANT92}.
+
 Already in the inital paper describing OBDDs, the importance of the variable
@@
 Nonetheless representing Boolean functions using OBDDs has some drawbacks. Classes
-of Boolean functions exist for which the size of the graph grows exponentially,
-e.g. integer multiplication~\cite{BRYANT86, BRYANT91, WOELFEL01}.
+of Boolean functions exist for which the size of the graph grows exponentially
+independent of the variable ordering, e.g. integer
+multiplication~\cite{BRYANT86, BRYANT91, WOELFEL01}.
 
-These limitations of OBDDs motivated researchers to introduce other
+These limitations of OBDDs motivated researchers to introduce heuristics and
+other techniques like \textit{dynamic variable ordering}~\cite{RUDELL93} to solve
+the problem of variable ordering~\cite{BOLLIG96}. Furthermore, alternative
 representations which e.g. relax the ordering~\cite{BRYANT95} or change the
-interpretation of nodes~\cite{BRYANT95}~\cite{ANDERSEN97}. Such representations
-may be beneficial for some classes of Boolean functions but mostly suffer from
-losing the canonical form and hence the desirable properties OBDDs are famous for.
+interpretation of nodes~\cite{BRYANT95, ANDERSEN97} were developed. Such
+representations may be beneficial for some classes of Boolean functions but mostly
+suffer from losing the canonical form and hence the desirable properties OBDDs
+are famous for.
 
@@
 discusses alternative representations to possibly overcome the limitations of
-the OBDD representation.
-\tbr
+the OBDD representation. \tbr
 
@@
 
-\section{Citations (temporary)}
-
-\noindent
-\par
-BRACE90\cite{BRACE90}, BRYANT92\cite{BRYANT92},
-\par
-RUDELL93\cite{RUDELL93}, BOLLIG96\cite{BOLLIG96}
-
 % balance columns on last page

### 2015-09-08 paper.tex
Add BDD description

@@
 
-\tbd
+\textit{Binary decision diagrams} (\textit{BDDs}) have been well studied since
+Lee~\cite{LEE59} and especially Akers~\cite{AKERS78} introduced them. BDDs
+represent a Boolean function $f$ as an acyclic, directed graph (DAG). Nodes
+represent variables and have one incoming and two outgoing edges, each of which
+denotes the variable evaluates to 0 (left) or 1 (right), respectively. There exist
+two \textit{terminal} nodes, denoted \textbf{0} (\textbf{False}) and \textbf{1}
+(\textbf{True}), which represent the result of $f$. To evaluate $f$ for a given
+variable assignment, one just follows the path for the respective variable values
+and the terminal nodes one exists is the result.
+\tbr
 

### 2015-09-08 paper.tex
Fix wording and typo

@@
 variable assignment, one just follows the path for the respective variable values
-and the terminal nodes one exists is the result.
+and the terminal node at the end is the result.
 \tbr

### 2015-09-08 paper.tex
Fix paragraph content

@@
 represent variables and have one incoming and two outgoing edges, each of which
-denotes the variable evaluates to 0 (left) or 1 (right), respectively. There exist
-two \textit{terminal} nodes, denoted \textbf{0} (\textbf{False}) and \textbf{1}
-(\textbf{True}), which represent the result of $f$. To evaluate $f$ for a given
-variable assignment, one just follows the path for the respective variable values
-and the terminal node at the end is the result.
+denotes the variable evaluates to 0 (left) or 1 (right), respectively. Whenever
+the result of $f$ is fixed, an edge points to one of the constants 0 or 1.
+To evaluate $f$ for a given variable assignment, one just follows the path for
+the respective variable values. As soon as the path ends up in a constant, i.e. 
+0 or 1, this constant value is the result of the variable assignment.
 \tbr

### 2015-09-08 figs/obdd.tex
Add OBDD example figure

\documentclass{standalone}

\usepackage{tikz}
\usetikzlibrary{backgrounds, positioning, decorations.pathreplacing, calc, fit}
\usetikzlibrary{shapes}

\begin{document}

\begin{tikzpicture}
    \tikzstyle{inner} = [draw, circle, thick];
    \tikzstyle{terminal} = [draw, rectangle, thick];

    \node[inner] (A) at (2, 0) {\huge{A}};
    \node[inner, below left = 1 and 2 of A] (Bl) {\huge{B}};
    \node[inner, below right = 1 and 2 of A] (Br) {\huge{B}};
    \node[inner, below = 3 of A] (C) {\huge{C}};

    \node[terminal, below left = 1.5 and 1 of C] (0) {\huge{0}};
    \node[terminal, below right = 1.5 and 1 of C] (1) {\huge{1}};

    \draw[dashed] (A.south west) -- (Bl.north east);
    \draw (A.south east) -- (Br.north west);

    \draw[dashed] (Bl.south east) -- (C.north west);
    \draw (Bl.south) -- (0.north);

    \draw[dashed] (Br.south west) -- (C.north east);
    \draw (Br.south) -- (1.north);

    \draw[dashed] (C.south west) -- (0.north);
    \draw (C.south east) -- (1.north);
\end{tikzpicture}

\end{document}

### 2015-09-08 paper.tex
Add BDD and OBDD example figures

@@
 \documentclass{vldb}
+
+%%% Packages
 \usepackage{graphicx}
 \usepackage{balance}
+\usepackage{tikz}
+\usepackage{standalone}
+\usepackage{subfigure}
 
+%%% TikZ libraries
+\usetikzlibrary{backgrounds, positioning, decorations.pathreplacing, calc, fit}
+\usetikzlibrary{shapes}
+
+%%% Macros
 \newcommand{\tbr}{\textbf{[TO BE REVISED]}}
@@
 
-% title
+% Title
 \title{A Survey on Ordered Binary Decision Diagrams}
 
-% author(s)
+% Author(s)
 \numberofauthors{1}
@@
 denotes the variable evaluates to 0 (left) or 1 (right), respectively. Whenever
-the result of $f$ is fixed, an edge points to one of the constants 0 or 1.
-To evaluate $f$ for a given variable assignment, one just follows the path for
-the respective variable values. As soon as the path ends up in a constant, i.e. 
-0 or 1, this constant value is the result of the variable assignment.
+the result of $f$ is fixed, an edge points to one of the constant functions
+\textbf{0} or \textbf{1}. To evaluate $f$ for a given variable assignment, one
+just follows the path for the respective variable values. As soon as the path
+ends up in a constant, i.e. \textbf{0} or \textbf{1}, this is the result of the
+variable assignment. Akers~\cite{AKERS78} already provided some techniques to
+simplify the BDD representation such as combining equivalent nodes.
+
+\begin{figure}[ht]
+    \centering
+    \subfigure[BDD \newline $f=(A \land B) \lor (B \land C)$.]{
+        \includestandalone[height = 3cm]{figs/bdd}
+        \label{fig:bdd}
+    }
+    \hskip1cm
+    \subfigure[OBDD \newline $f=(A \land B) \lor (B \land C)$.]{
+        \includestandalone[height = 3cm]{figs/obdd}
+        \label{fig:obdd}
+    }
+\end{figure}
+
 \tbr

### 2015-09-10 paper.tex
Finish representation basics introduction

@@
 ends up in a constant, i.e. \textbf{0} or \textbf{1}, this is the result of the
-variable assignment. Akers~\cite{AKERS78} already provided some techniques to
-simplify the BDD representation such as combining equivalent nodes.
+corresponding variable assignment. Akers~\cite{AKERS78} already provided some
+techniques to simplify the BDD representation such as combining equivalent nodes.
+
+Figure~\ref{fig:bdd} shows an example of a BDD representing the Boolean function
+$f=(A \land B) \lor (B \land C)$. To simplify the graphs, the diagram style of
+Bryant~\cite{BRYANT92} is reused, i.e. edge labels are skipped. Instead dashed
+and solid lines are used to represent the value of the decision variable to be 0
+and 1, respectively. This diagram style is used throughout this paper.
 
@@
     \centering
-    \subfigure[BDD \newline $f=(A \land B) \lor (B \land C)$.]{
+    \subfigure[BDD]{
         \includestandalone[height = 3cm]{figs/bdd}
@@
     \hskip1cm
-    \subfigure[OBDD \newline $f=(A \land B) \lor (B \land C)$.]{
+    \subfigure[OBDD]{
         \includestandalone[height = 3cm]{figs/obdd}
@@
     }
+    \caption{BDD and OBDD of $f=(A \land B) \lor (B \land C)$.}
 \end{figure}
+
+However, Bryant~\cite{BRYANT86} introduced some restrictions concerning the
+ordering of the so-called \textit{decision variables} represented by the vertices
+of the BDD. Furthermore, Bryant\cite{BRYANT86} was able to derive a reduced
+representation of BDDs providing the DAG in a canonical form. The following two
+subsections cover the details.
 
@@
 
-\subsection{Reducing and Ordering BDDs}
-\label{subsec:reducing-and-ordering-bdds}
+\subsection{From BDDs to OBDDs}
+\label{subsec:from-bdds-to-obdds}
 
@@
 
-\tbd
+\begin{figure}[ht]
+    \centering
+    \includestandalone[height = 3cm]{figs/min-obdd}
+    \label{fig:min-obdd}
+    \caption{Minimized OBDD (optimal).}
+\end{figure}
+\tbr
 

### 2015-09-10 paper.tex
Fix wording

@@
 However, Bryant~\cite{BRYANT86} introduced some restrictions concerning the
-ordering of the so-called \textit{decision variables} represented by the vertices
+ordering of the so-called \textit{decision variables} represented by the nodes
 of the BDD. Furthermore, Bryant\cite{BRYANT86} was able to derive a reduced

### 2015-09-10 paper.tex
Fix formulations and add ROBDD term

@@
 of the BDD. Furthermore, Bryant\cite{BRYANT86} was able to derive a reduced
-representation of BDDs providing the DAG in a canonical form. The following two
-subsections cover the details.
-
-\tbr
+representation of BDDs providing the DAG in a canonical form: the
+\textit{Reduced Ordered Binary Decision Diagram} (\textit{ROBDD}; mostly referred
+to as \textit{OBDD}, even by Bryant himself). The following section covers the
+details of the OBDD representation. \tbr
 

### 2015-09-10 paper.tex
Add notice that OBDD and ROBDD are used synonymously
Add figure reference and fix wording

@@
 \textit{Reduced Ordered Binary Decision Diagram} (\textit{ROBDD}; mostly referred
-to as \textit{OBDD}, even by Bryant himself). The following section covers the
-details of the OBDD representation. \tbr
+to as \textit{OBDD}, even by Bryant himself). In this paper, the terms ROBDD and
+OBDD are used synonymously.
+
+Figure~\ref{fig:obdd} shows one possible ROBDD representation of
+$f=(A \land B) \lor (B \land C)$. The following section covers the details of how
+to derive the OBDD representation from a BDD. \tbr
 

### 2015-09-10 paper.tex
Remove duplicate function definition
Add first two paragraphs to From BDDs to OBDDs section

@@
 Figure~\ref{fig:obdd} shows one possible ROBDD representation of
-$f=(A \land B) \lor (B \land C)$. The following section covers the details of how
+$f$. The following section covers the details of how
 to derive the OBDD representation from a BDD. \tbr
@@
 
-\tbd
+As the name suggests, \textit{Reduced} Ordered Binary Decision Diagrams are a 
+minimized representation of BDDs. Furthermore, the decision variables are ordered
+consistently. More formally, for any node $\upsilon$ and their respective children
+$\omega_1$ and $\omega_2$, a total ordering is imposed, i.e.
+$var(\upsilon) < var(\omega_1)$ and $var(\upsilon) < var(\omega_2)$, respectively,
+where $var(x)$ denotes the label of node $x$.
+
+This \textit{variable ordering} affects the efficiency of the symbolic
+manipulation because different variable orderings, in general, produce different
+OBDDs. In the worst case, this will result in exponential complexity for bad
+orderings. Nevertheless, as will be described in Section~\ref{sec:limitations},
+there exist classes of functions for which the OBDD complexity is exponential
+regardless of which variable ordering is chosen.
+
+\tbr
 

### 2015-09-10 paper.tex
Add citations
Add paragraphs (definitions)

@@
 $var(\upsilon) < var(\omega_1)$ and $var(\upsilon) < var(\omega_2)$, respectively,
-where $var(x)$ denotes the label of node $x$.
+where $var(x)$ denotes the label of node $x$~\cite{BRYANT86, BRYANT92}.
 
@@
 regardless of which variable ordering is chosen.
+
+Before discussing the reduction procedure, some definitions are necessary. Here
+the original definitions of Bryant~\cite{BRYANT86, BRYANT91} are reused.
+
+A Boolean function can be represented as an OBDD, i.e. a rooted, directed,
+acyclic graph (DAG) with a set of nodes $V$ as well as a set of edges $E$.
+
+There exist \textit{nonterminal} and \textit{terminal} nodes. Each
+\textit{nonterminal} node $n$ has its own label $var(n)$ as well as two edges,
+connecting it to its left $lo(n)\in V$ and right child $hi(n)\in V$, respectively. 
+Each \textit{terminal} node $t$ has a unique value $value(t)$, i.e. $value(t)$
+is a constant function (for Boolean functions \textbf{0} or \textbf{1}). Hence,
+for each constant function only a single terminal node exists.
+
+The following property has to hold for any nonterminal node $n$ in order to
+satisfy the ordering requirement of an OBDD:
+$var(n) < var(lo(n)) \land var(n) < var(hi(n))$.
+In other words, for any path starting at the root of an OBDD, if a path is
+followed until a terminal node is reached, the nonterminal nodes are traversed
+strictly in the variable ordering.
 

### 2015-09-10 paper.tex
Add additional paragraphs (definitions)

@@
 
+\paragraph*{Reduction procedure}
+
+Bryant~\cite{BRYANT86, BRYANT92} described three transformation rules to reduce
+a graph such that it still represents the same function as before.
+
+For each constant function, there exists only a single terminal node representing
+it, e.g. for Boolean functions, only two terminal nodes exist, representing
+\textbf{0} and \textbf{1}, respectively. Hence, when transforming a given BDD,
+eliminate all but one terminal node for each constant function and redirect all
+edges accordingly, i.e. to the one remaining node for each constant function.
+
+There may be duplicate nonterminal nodes present in a given BDD. Formally, a node
+$d$ is a duplicate if $var(d) = var(n) \land lo(d) = lo(n) \land hi(d) = hi(n)$
+holds for any other node $n$. If this is the case, $d$ can be eliminated and all
+edges are redirected to $n$.
+
+The third and thus last transformation rule deals with completely redundant 
+nonterminal nodes. Intuitively, this rule should be straightforward: any node
+having the same child on all outgoing edges is redundant because the actual value
+of the variable does not change the result of the function. More formally, a
+nonterminal node $n$ is redundant if $lo(n) = hi(n)$. Then, all incoming edges
+of $n$ can be redirected to $lo(n)$ and $n$ can be deleted afterwards.
+
 \tbr

### 2015-09-10 figs/obdd.tex
Fix line width

@@
 
-    \draw[dashed] (A.south west) -- (Bl.north east);
-    \draw (A.south east) -- (Br.north west);
+    \draw[thick, dashed] (A.south west) -- (Bl.north east);
+    \draw[thick] (A.south east) -- (Br.north west);
 
-    \draw[dashed] (Bl.south east) -- (C.north west);
-    \draw (Bl.south) -- (0.north);
+    \draw[thick, dashed] (Bl.south east) -- (C.north west);
+    \draw[thick] (Bl.south) -- (0.north);
 
-    \draw[dashed] (Br.south west) -- (C.north east);
-    \draw (Br.south) -- (1.north);
+    \draw[thick, dashed] (Br.south west) -- (C.north east);
+    \draw[thick] (Br.south) -- (1.north);
 
-    \draw[dashed] (C.south west) -- (0.north);
-    \draw (C.south east) -- (1.north);
+    \draw[thick, dashed] (C.south west) -- (0.north);
+    \draw[thick] (C.south east) -- (1.north);
 \end{tikzpicture}

### 2015-09-10 paper.tex
Add workaround to get line break after paragraph
Reformat
Add figures and references

@@
     }
+    \label{fig:bdd-and-obdd}
     \caption{BDD and OBDD of $f=(A \land B) \lor (B \land C)$.}
@@
 \paragraph*{Reduction procedure}
+\mbox{} % workaround to get line break after paragraph
 
@@
 
-The third and thus last transformation rule deals with completely redundant 
-nonterminal nodes. Intuitively, this rule should be straightforward: any node
-having the same child on all outgoing edges is redundant because the actual value
-of the variable does not change the result of the function. More formally, a
-nonterminal node $n$ is redundant if $lo(n) = hi(n)$. Then, all incoming edges
-of $n$ can be redirected to $lo(n)$ and $n$ can be deleted afterwards.
+The third and last transformation rule deals with completely redundant nonterminal
+nodes. Intuitively, this rule should be straightforward: any node having the same
+child on all outgoing edges is redundant because the actual value of the variable
+does not change the result of the function. More formally, a nonterminal node $n$
+is redundant if and only if $lo(n) = hi(n)$. Then, all incoming edges of $n$ can
+be redirected to $lo(n)$ and $n$ can be deleted afterwards.
 
-\tbr
+\begin{figure}[ht]
+    \centering
+    \subfigure[]{
+        \includestandalone[height = 3cm]{figs/redundant-nonterminals-before}
+        \label{fig:redundant-nonterminals-before}
+    }
+    \hskip1cm
+    \subfigure[]{
+        \includestandalone[height = 3cm]{figs/redundant-nonterminals-after}
+        \label{fig:redundant-nonterminals-after}
+    }
+    \label{fig:redundant-nonterminals}
+    \caption{Eliminating redundant nonterminal nodes.}
+\end{figure}
+
+Figure~\ref{fig:obdd} shows the result of applying the first two transformation
+rules to the BDD shown in Figure~\ref{fig:bdd}. The third rule is shown
+in Figure~\ref{fig:redundant-nonterminals}.
+Figure~\ref{fig:redundant-nonterminals-before} shows a BDD with redundant
+nonterminal nodes (both \textbf{B}'s, in essence) and
+Figure~\ref{fig:redundant-nonterminals-after} shows the equivalent OBDD without
+redundant nonterminals. \tbr
 

### 2015-09-10 paper.tex
Fix figure references

@@
 
-Figure~\ref{fig:bdd} shows an example of a BDD representing the Boolean function
+Figure~\ref{subfig:bdd} shows an example of a BDD representing the Boolean function
 $f=(A \land B) \lor (B \land C)$. To simplify the graphs, the diagram style of
@@
         \includestandalone[height = 3cm]{figs/bdd}
-        \label{fig:bdd}
+        \label{subfig:bdd}
     }
@@
         \includestandalone[height = 3cm]{figs/obdd}
-        \label{fig:obdd}
+        \label{subfig:obdd}
     }
+    \caption{BDD and OBDD of $f=(A \land B) \lor (B \land C)$.}
     \label{fig:bdd-and-obdd}
-    \caption{BDD and OBDD of $f=(A \land B) \lor (B \land C)$.}
 \end{figure}
@@
 
-Figure~\ref{fig:obdd} shows one possible ROBDD representation of
+Figure~\ref{subfig:obdd} shows one possible ROBDD representation of
 $f$. The following section covers the details of how
@@
         \includestandalone[height = 3cm]{figs/redundant-nonterminals-before}
-        \label{fig:redundant-nonterminals-before}
+        \label{subfig:redundant-nonterminals-before}
     }
@@
         \includestandalone[height = 3cm]{figs/redundant-nonterminals-after}
-        \label{fig:redundant-nonterminals-after}
+        \label{subfig:redundant-nonterminals-after}
     }
+    \caption{Eliminating redundant nonterminal nodes.}
     \label{fig:redundant-nonterminals}
-    \caption{Eliminating redundant nonterminal nodes.}
 \end{figure}
 
-Figure~\ref{fig:obdd} shows the result of applying the first two transformation
-rules to the BDD shown in Figure~\ref{fig:bdd}. The third rule is shown
+Figure~\ref{subfig:obdd} shows the result of applying the first two transformation
+rules to the BDD shown in Figure~\ref{subfig:bdd}. The third rule is shown
 in Figure~\ref{fig:redundant-nonterminals}.
-Figure~\ref{fig:redundant-nonterminals-before} shows a BDD with redundant
+Figure~\ref{subfig:redundant-nonterminals-before} shows a BDD with redundant
 nonterminal nodes (both \textbf{B}'s, in essence) and
-Figure~\ref{fig:redundant-nonterminals-after} shows the equivalent OBDD without
+Figure~\ref{subfig:redundant-nonterminals-after} shows the equivalent OBDD without
 redundant nonterminals. \tbr

### 2015-09-10 paper.tex
Fix wording

@@
 
-Figure~\ref{subfig:obdd} shows the result of applying the first two transformation
-rules to the BDD shown in Figure~\ref{subfig:bdd}. The third rule is shown
-in Figure~\ref{fig:redundant-nonterminals}.
+Figure~\ref{fig:bdd-and-obdd} shows the result, \ref{subfig:obdd}, of applying
+the first two transformation rules to the BDD shown in \ref{subfig:bdd}.
+The third rule is shown in Figure~\ref{fig:redundant-nonterminals}.
 Figure~\ref{subfig:redundant-nonterminals-before} shows a BDD with redundant
-nonterminal nodes (both \textbf{B}'s, in essence) and
-Figure~\ref{subfig:redundant-nonterminals-after} shows the equivalent OBDD without
-redundant nonterminals. \tbr
+nonterminal nodes (both B's, in essence) and
+Figure~\ref{subfig:redundant-nonterminals-after} shows the equivalent OBDD
+without redundant nonterminals. \tbr
 

### 2015-09-11 paper.tex
Fix wording

@@
 the first two transformation rules to the BDD shown in \ref{subfig:bdd}.
-The third rule is shown in Figure~\ref{fig:redundant-nonterminals}.
-Figure~\ref{subfig:redundant-nonterminals-before} shows a BDD with redundant
+The third rule is shown in Figure~\ref{fig:redundant-nonterminals}:
+\ref{subfig:redundant-nonterminals-before} shows a BDD with redundant
 nonterminal nodes (both B's, in essence) and
-Figure~\ref{subfig:redundant-nonterminals-after} shows the equivalent OBDD
-without redundant nonterminals. \tbr
+\ref{subfig:redundant-nonterminals-after} shows the equivalent OBDD without
+redundant nonterminals. \tbr
 

### 2015-09-11 paper.tex
Finish first version of Representation basics

@@
 \ref{subfig:redundant-nonterminals-after} shows the equivalent OBDD without
-redundant nonterminals. \tbr
+redundant nonterminals.
+
+Bryant already formally described an algorithm to transform an arbitrary graph
+into a reduced one~\cite[p. 683]{BRYANT86}.
+
+OBDDs have one big advantage over other representations: their representation is
+\textit{canonical}. As a consequence, important tests can be performed on an OBDD
+in constant time, i.e. satisfiability and tautology tests. Moreover, OBDDs enable
+other tests, such as variable independence and functional equivalence tests, to
+be performed efficiently. \tbr
 

### 2015-09-11 paper.tex
Fix heading
Add Operations intro

@@
 
-\section{Efficient Operations}
-\label{sec:efficient-operations}
+\section{Operations}
+\label{sec:operations}
 
-\tbd
+Bryant's~\cite{BRYANT86, BRYANT92} main focus was to define algorithms performing
+symbolic operations on graphs representing Boolean functions. The most basic
+operation is \textit{Apply} which applys a given operation $op$, e.g. \texttt{AND},
+to two given functions $f_1$ and $f_2$, and returns the result of $f_1{ }op{ }f_2$.
+Another operation is \textit{Restrict}, restricting a given variable $x$ to a
+given (constant) value $c$, denoted $f_{x\leftarrow c}$. Furthermore, the
+operations \textit{Compose} and \textit{Satisfy} can be defined as graph
+algorithms. \tbr
 
@@
 \end{figure}
-\tbr
+\tbd
 

### 2015-09-11 paper.tex
Start apply operation description

@@
 operations \textit{Compose} and \textit{Satisfy} can be defined as graph
-algorithms. \tbr
+algorithms.
+
+The following paragraphs summarize those operations and their properties briefly.
+
+\paragraph*{Apply}
+\mbox{} % workaround to get line break after paragraph
+
+Applies an operation $op$ to two given functions $f_1$ and $f_2$ and returns the
+resulting function $f_1{ }op{ }f_2$. This is the most intuitive function one would
+expect when dealing with Boolean functions and it has some interesting properties
+too, i.e. the complement of $f_1$ can be computed by setting $f_2$ to \textbf{1}
+and $op$ to $\oplus$ (\texttt{XOR}). Moreover, the already mentioned functional
+equivalence can be tested by applying $\overline{\oplus}$ (\texttt{XNOR}) to
+$f_1$ and $f_2$ for all relevant cases. Therefore Bryant~\cite{BRYANT92} defined
+a function $d(x_1, \ldots, x_n)$ for \textit{Don't cares}:
+
+$d(x_1, \ldots, x_n)=
+\begin{cases}
+    1, & \text{if values are not relevant} \\
+    0, & otherwise
+\end{cases}
+$
+
+Using this function, one check if $f_1$ and $f_2$ are functionally equivalent by
+performing $(f_1\overline{\oplus}f_2)\lor d$. \tbr
+
+\paragraph*{Restrict}
+\mbox{} % workaround to get line break after paragraph
+
+\tbd
+
+\paragraph*{Compose}
+\mbox{} % workaround to get line break after paragraph
+
+\tbd
+
+\paragraph*{Satisfy}
+\mbox{} % workaround to get line break after paragraph
+
+\tbd
 

### 2015-09-11 paper.tex
Finish Apply operation

@@
 operation is \textit{Apply} which applys a given operation $op$, e.g. \texttt{AND},
-to two given functions $f_1$ and $f_2$, and returns the result of $f_1{ }op{ }f_2$.
+to two given functions $f_1$ and $f_2$, and returns the result of $f_1\;op\;f_2$.
 Another operation is \textit{Restrict}, restricting a given variable $x$ to a
-given (constant) value $c$, denoted $f_{x\leftarrow c}$. Furthermore, the
+given (constant) value $c$, denoted $f|_{x\leftarrow c}$. Furthermore, the
 operations \textit{Compose} and \textit{Satisfy} can be defined as graph
@@
 The following paragraphs summarize those operations and their properties briefly.
+Implementation details are covered in Section~\ref{sec:implementation-aspects}.
+
+\paragraph*{Restrict}
+\mbox{} % workaround to get line break after paragraph
+
+\tbd
 
@@
 Applies an operation $op$ to two given functions $f_1$ and $f_2$ and returns the
-resulting function $f_1{ }op{ }f_2$. This is the most intuitive function one would
+resulting function $f_1\;op\;f_2$. This is the most intuitive function one would
 expect when dealing with Boolean functions and it has some interesting properties
@@
 equivalence can be tested by applying $\overline{\oplus}$ (\texttt{XNOR}) to
-$f_1$ and $f_2$ for all relevant cases. Therefore Bryant~\cite{BRYANT92} defined
+$f_1$ and $f_2$ for all relevant cases. Therefore, Bryant~\cite{BRYANT92} defined
 a function $d(x_1, \ldots, x_n)$ for \textit{Don't cares}:
@@
 Using this function, one check if $f_1$ and $f_2$ are functionally equivalent by
-performing $(f_1\overline{\oplus}f_2)\lor d$. \tbr
+computing $(f_1\overline{\oplus}f_2)\lor d$. If the result is \textbf{1}, $f_1$
+and $f_2$ are functionally equivalent.
 
-\paragraph*{Restrict}
-\mbox{} % workaround to get line break after paragraph
+\textit{Apply} traverses the given graph depth-first, creating the resultant OBDD
+step by step. Starting from the roots $n_1$ and $n_2$, new nodes are created for
+each branching point of the graphs of $f_1$ and $f_2$, respectively.
 
-\tbd
+\textit{Shannon's expansion} forms the basis of the recursive
+algorithm~\cite{BRYANT86}:
+\begin{center}
+$
+f_1\;op\;f_2 =
+\overline{x} \cdot \left(f_1|_{x\leftarrow 0}\;op\;f_2|_{x\leftarrow 0}\right)
++ x \cdot \left(f_1|_{x\leftarrow 1}\;op\;f_2|_{x\leftarrow 1}\right)
+$
+\end{center}
+
+The recursive procedure then has three different cases. Note that a node is only
+created if no equivalent node is already present in the resultant graph.
+\begin{enumerate}
+    \item{
+        If $n_1$ and $n_2$ are both terminal nodes, then a terminal node
+        representing the value $value(n_1)\;op\;value(n_2)$ is added to the
+        resultant graph and the recursion terminates.
+    }
+    \item{
+        If either $n_1$ or $n_2$ is a nonterminal node, then:
+        \begin{enumerate}
+            \item{
+                If $var(n_1) = var(n_2)$, a node having $var(n_1)$ is created and
+                the procedure is applied recursively on $(lo(n_1), lo(n_2))$ and
+                $(hi(n_1), hi(n_2))$.
+            }
+            \item{
+                If $var(n_1)$ is a nonterminal and $n_2$ is eiter a terminal node
+                or $var(n_1) < var(n_2)$, a node having $var(n_1)$ is created and
+                the procedure is applied recursively on $(lo(n_1), n_2)$ and
+                $(hi(n_1), n_2)$.
+            }
+        \end{enumerate}
+    }
+\end{enumerate}
+
+By applying the three transformation rules of the reduction procedure in each
+step, \textit{Apply} generates a reduced graph immediately~\cite{BRYANT92}.
+
+There are some refinements when it comes to implementation. Those are covered in
+Section~\ref{sec:implementation-aspects} as well as the original paper of
+Bryant~\cite{BRYANT86, BRYANT92}. \tbr
 

### 2015-09-11 paper.tex
Add function argument omittance notice
Finish Restrict operation; add time complexity to Apply operation

@@
 the corresponding papers easily (references are provided for each algorithm).
-Section~\ref{sec:implementation-aspects} describes aspects one has to take into
+Section~\ref{sec:implementational-aspects} describes aspects one has to take into
 account when implementing an OBDD package, such as what data structures to use
@@
 Lee~\cite{LEE59} and especially Akers~\cite{AKERS78} introduced them. BDDs
-represent a Boolean function $f$ as an acyclic, directed graph (DAG). Nodes
-represent variables and have one incoming and two outgoing edges, each of which
-denotes the variable evaluates to 0 (left) or 1 (right), respectively. Whenever
-the result of $f$ is fixed, an edge points to one of the constant functions
-\textbf{0} or \textbf{1}. To evaluate $f$ for a given variable assignment, one
-just follows the path for the respective variable values. As soon as the path
-ends up in a constant, i.e. \textbf{0} or \textbf{1}, this is the result of the
-corresponding variable assignment. Akers~\cite{AKERS78} already provided some
-techniques to simplify the BDD representation such as combining equivalent nodes.
+represent a Boolean function $f$ as an acyclic, directed graph (DAG). The function
+arguments may be omitted in some cases to keep the formulas simple, i.e. $f$
+denotes $f(x_1, \ldots, x_2)$. Nodes represent variables and have one incoming
+and two outgoing edges, each of which denotes the variable evaluates to 0 (left)
+or 1 (right), respectively. Whenever the result of $f$ is fixed, an edge points
+to one of the constant functions \textbf{0} or \textbf{1}. To evaluate $f$ for a
+given variable assignment, one just follows the path for the respective variable
+values. As soon as the path ends up in a constant, i.e. \textbf{0} or \textbf{1},
+this is the result of the corresponding variable assignment. Akers~\cite{AKERS78}
+already provided some techniques to simplify the BDD representation such as
+combining equivalent nodes.
 
@@
 The following paragraphs summarize those operations and their properties briefly.
-Implementation details are covered in Section~\ref{sec:implementation-aspects}.
+Implementation details are covered in Section~\ref{sec:implementational-aspects}.
 
@@
 
-\tbd
+Although \textit{Apply} is the core operation, the \textit{Restrict} operation is
+discussed first, mainly because it is used within \textit{Apply}. Besides,
+\textit{Restrict} is straightforward and easy to understand.
+
+A variable $x$, argument of a function $f$, can be restricted to a given value
+$c$, denoted by $f|_{x\leftarrow c}$, as follows.
+
+Starting from the root $n$ of the OBDD representation of $f$ and traversing the
+OBDD depth-first, every node is checked for $var(n) = x$. This is done recursively
+on the given graph, generating the resultant graph This node represents
+the variable to be restricted. Now, there are two cases for $c$ to distinguish:
+\begin{enumerate}
+    \item{$c=0$: The incoming edge of $n$ is redirected to $lo(n)$.}
+    \item{$c=1$: The incoming edge of $n$ is redirected to $hi(n)$.}
+\end{enumerate}
+
+By applying the three transformation rules of the reduction procedure in each
+step, \textit{Restrict} generates a reduced graph immediately~\cite{BRYANT92}.
+
+This procedure has linear time complexity in the number of nodes, i.e. $O(|V|)$,
+assuming an efficient implementation (see
+Section~\ref{sec:implementational-aspects} for details)~\cite{BRYANT92}. 
+\tbr
 
@@
 
-By applying the three transformation rules of the reduction procedure in each
-step, \textit{Apply} generates a reduced graph immediately~\cite{BRYANT92}.
+To get a reduced graph immediately, the three transformation rules are utilized
+again for each step~\cite{BRYANT92}.
 
 There are some refinements when it comes to implementation. Those are covered in
-Section~\ref{sec:implementation-aspects} as well as the original paper of
-Bryant~\cite{BRYANT86, BRYANT92}. \tbr
+Section~\ref{sec:implementational-aspects} as well as the original paper of
+Bryant~\cite{BRYANT86, BRYANT92}.  However, assuming an efficient implementation,
+the time complexity can be bound to
+$O\left(|V_{f_1}| \cdot |V_{f_2}|\right)$~\cite{BRYANT86, BRYANT92}. \tbr
 
@@
 \section{Implementational Aspects}
-\label{sec:implementation-aspects}
+\label{sec:implementational-aspects}
 

### 2015-09-11 paper.tex
Add reference to Apply pseudocode
Add author to Apply pseudocode reference
Start Compose operation
Fix formula notation to be consistent
Fix node notation to be consistent

@@
 minimized representation of BDDs. Furthermore, the decision variables are ordered
-consistently. More formally, for any node $\upsilon$ and their respective children
-$\omega_1$ and $\omega_2$, a total ordering is imposed, i.e.
-$var(\upsilon) < var(\omega_1)$ and $var(\upsilon) < var(\omega_2)$, respectively,
-where $var(x)$ denotes the label of node $x$~\cite{BRYANT86, BRYANT92}.
+consistently. More formally, for any node $n$ and their respective children $m_1$
+and $m_2$, a total ordering is imposed, i.e. $var(n) < var(m_1)$ and
+$var(n) < var(m_2)$, respectively, where $var(x)$ denotes the label of node
+$x$~\cite{BRYANT86, BRYANT92}.
 
@@
 f_1\;op\;f_2 =
-\overline{x} \cdot \left(f_1|_{x\leftarrow 0}\;op\;f_2|_{x\leftarrow 0}\right)
-+ x \cdot \left(f_1|_{x\leftarrow 1}\;op\;f_2|_{x\leftarrow 1}\right)
+\overline{x} \land \left(f_1|_{x\leftarrow 0}\;op\;f_2|_{x\leftarrow 0}\right)
+\lor x \land \left(f_1|_{x\leftarrow 1}\;op\;f_2|_{x\leftarrow 1}\right)
 $
@@
 the time complexity can be bound to
-$O\left(|V_{f_1}| \cdot |V_{f_2}|\right)$~\cite{BRYANT86, BRYANT92}. \tbr
+$O\left(|V_{f_1}| \cdot |V_{f_2}|\right)$~\cite{BRYANT86, BRYANT92}.
+
+The full pseudocode defined by Bryant can be found in~\cite[p. 685]{BRYANT86}. \tbr
 
@@
 
-\tbd
+This operation is merely based upon calls to \textit{Apply} and \textit{Restrict}.
+Utilizing \textit{Shannon's expansion}, Bryant~\cite{BRYANT86} formulated the
+composition of two functions $f_1$ and $f_2$ as
+\begin{center}
+$
+f_1|_{x=f_2}=
+f_2 \land f_1|_{x_i=1} \lor \overline{f_2} \land f_1|_{x_i=0}
+$
+\end{center}\tbr
 

### 2015-09-11 paper.tex
Add content Compose operation

@@
 $
-\end{center}\tbr
+\end{center}
+
+However, this would result in an algorithm having a time complexity of 
+$O\left(|V_{f_1}|^2 \cdot |V_{f_2}|^2\right)$. Therefore, Bryant~\cite{BRYANT86}
+already proposed to extend the \textit{Apply} operation to support ternary
+operations and then use the equivalent \textit{if-then-else} (\textit{ITE})
+representation:
+\begin{center}
+$
+ITE\left(f_2, f_1|_{x_i=1}, f_1|_{x_i=0}\right) \equiv
+f_2 \land f_1|_{x_i=1} \lor \overline{f_2} \land f_1|_{x_i=0}
+$
+\end{center}
+
+\noindent
+The time complexity then improves to be
+$O\left(|V_{f_1}|^2 \cdot |V_{f_2}|\right)$~\cite{BRYANT86}. \tbr
 

### 2015-09-11 paper.tex
Add \noindent to paragraphs following formulas

@@
 
+\noindent
 Using this function, one check if $f_1$ and $f_2$ are functionally equivalent by
@@
 
+\noindent
 The recursive procedure then has three different cases. Note that a node is only

### 2015-09-11 paper.tex
Finish Compose operation
Restructure
Finish Satisfy operation

@@
 The time complexity then improves to be
-$O\left(|V_{f_1}|^2 \cdot |V_{f_2}|\right)$~\cite{BRYANT86}. \tbr
+$O\left(|V_{f_1}|^2 \cdot |V_{f_2}|\right)$~\cite{BRYANT86}.
+~\cite[p. 686]{BRYANT86} lists Bryant's full pseudocode.
+
+Bryant~\cite{BRYANT86} also described another, even more efficient way to compose
+two function graphs. But this approach only applicable under restricted conditions.
+\tbr
 
@@
 
-\tbd
+In essence, Bryant~\cite{BRYANT86} described two operations, i.e.
+\textit{Satisfy-one} and \textit{Satisfy-all}, both operating on the satisfying
+set. As the name suggests, \textit{Satisfy-one} selects a single element and
+\textit{Satisfy-all} selects all elements from the satisfying set.
+
+\textit{Satisfy-one} traverses the OBDD depth-first and uses backtracking until
+a terminal node of the constant function \textbf{1} is encountered. If such a
+node is found, the corresponding Boolean function is satisfiable and true is
+returned. Furthermore, an array representing the satisfying variable assignment
+is assigned. The time complexity for a reduced graph is linear, i.e.
+$O(|V_f|)$~\cite{BRYANT86}.
+
+\textit{Satisfy-all} also traverses the OBDD depth-first. Every time the terminal
+node having value \textbf{1} is traversed, it is printed. The recursive procedure
+terminates as soon as the terminal node having value \textbf{0} is reached. This
+results in a time complexity of $O(|V_f| \cdot |S_f|)$, whereas $S_f$ denotes the
+satisfying set~\cite{BRYANT86}. \tbr
 

### 2015-09-11 paper.tex
Fix wording
Add Implementational aspects intro

@@
 In essence, Bryant~\cite{BRYANT86} described two operations, i.e.
-\textit{Satisfy-one} and \textit{Satisfy-all}, both operating on the satisfying
+\textit{Satisfy-one} and \textit{Satisfy-all}, both concerned with the satisfying
 set. As the name suggests, \textit{Satisfy-one} selects a single element and
@@
 
-\tbd
+Bryant~\cite{BRYANT86} already observed the importance of utilizing efficient
+techniques to implement the operations. Computation applied to OBDDs has a highly
+dynamic character without specific memory access patterns~\cite{BRYANT92}.
+
+Hence, the following section cover important aspects to keep in mind when
+implementing an OBDD framework efficiently, i.e. data structures, how to deal with
+the variable ordering problem and \textit{garbage collection}. \tbr
 

### 2015-09-11 paper.tex
Fix wording

@@
 
-Hence, the following section cover important aspects to keep in mind when
-implementing an OBDD framework efficiently, i.e. data structures, how to deal with
-the variable ordering problem and \textit{garbage collection}. \tbr
+Hence, the following sections cover important aspects to keep in mind when
+implementing an OBDD framework efficiently, i.e. efficient data structures, how
+to possibly approach the variable ordering problem and \textit{garbage collection}.
+\tbr
 

### 2015-09-11 paper.tex
Add subsection Efficient Data Structures and Garbage Collection
Add hash table data structure

@@
 
+\subsection{Efficient Data Structures}
+\label{subsec:efficient-data-structures}
+
+\paragraph*{Hash Table}
+\mbox{} % workaround to get line break after paragraph
+
+Hash tables maintain \textit{key}-\textit{value}-pairs, returning the associated
+value for a given key in (amortized) constant time~\cite{BRACE90}.
+
+This is especially useful to refine \textit{Apply} in order to guarantee the time
+complexity of $O\left(|V_{f_1}| \cdot |V_{f_2}|\right)$. Using a hash table to
+match pairs of nodes ($n_1$, $n_2$) to a single node $m$, \textit{Apply} avoids
+making multiple recursive calls to equivalent pairs of nodes and thus avoids
+computing equivalent subgraphs multiple times. Therefore, before computing a
+subgraph, \textit{Apply} checks whether an entry is present for the given pair
+of nodes. If so, the associated value $m$ is used instead of computing the
+subgraph~\cite{BRYANT86, BRYANT92}.
+
+Moreover, hash tables are useful to generate a reduced graph directly from some
+operations by maintaing a hash table associating triples
+($var(n)$, $lo(n)$, $hi(n)$) with a single nonterminal node $n$. This is done for
+each nonterminal node $n$ generated while executing the
+operation~\cite{BRYANT92}.
+\tbr
+
 \subsection{Dynamic Variable Ordering}
@@
 \end{figure}
+\tbd
+
+\subsection{Garbage Collection}
+\label{subsec:garbage-collection}
+
 \tbd

### 2015-09-12 paper.tex
Add hash-based chaches to hash tables

@@
 
-\paragraph*{Hash Table}
+\paragraph*{Hash Tables}
 \mbox{} % workaround to get line break after paragraph
@@
 Hash tables maintain \textit{key}-\textit{value}-pairs, returning the associated
-value for a given key in (amortized) constant time~\cite{BRACE90}.
+value for a given key in (amortized) constant time. Collisions, i.e. equivalent
+value for different keys, are resolved by chaining all associated values in a 
+linked list~\cite{BRACE90}.
 
@@
 operation~\cite{BRYANT92}.
-\tbr
+
+There exists a special version of hash tables, so-called
+\textit{hash-based caches}, which do not resolve collisions, i.e. overwrite
+existing entries. \textit{Hash-based caches} issue a cache miss if a looked up
+element is not found. \tbr
 

### 2015-09-12 paper.tex
Add ITE operator to Implementational aspects intro
Add ITE operator subsection
Remove ITE emphasis
Finish hash-based caches

@@
 Hence, the following sections cover important aspects to keep in mind when
-implementing an OBDD framework efficiently, i.e. efficient data structures, how
-to possibly approach the variable ordering problem and \textit{garbage collection}.
-\tbr
+implementing an OBDD framework efficiently, i.e. efficient data structures, the
+ITE operator, how to possibly approach the variable ordering problem and
+\textit{garbage collection}. \tbr
 
@@
 \textit{hash-based caches}, which do not resolve collisions, i.e. overwrite
-existing entries. \textit{Hash-based caches} issue a cache miss if a looked up
-element is not found. \tbr
+existing entries, and issue a cache miss if a looked up element is not found.
+As will be seen, \textit{Hash-based caches} are useful to provide an efficient
+implementation of the ITE operator. \tbr
+
+\subsection{ITE Operator}
+\label{subsec:ite-operator}
+
+\tbd
 

### 2015-09-12 paper.tex
Start ITE operator
Add index to formulas

@@
 
-A variable $x$, argument of a function $f$, can be restricted to a given value
-$c$, denoted by $f|_{x\leftarrow c}$, as follows.
+A variable $x_i$, argument of a function $f$, can be restricted to a given value
+$c$, denoted by $f|_{x_i\leftarrow c}$, as follows.
 
 Starting from the root $n$ of the OBDD representation of $f$ and traversing the
-OBDD depth-first, every node is checked for $var(n) = x$. This is done recursively
+OBDD depth-first, every node is checked for $var(n) = x_i$. This is done recursively
 on the given graph, generating the resultant graph This node represents
@@
 $
-f_1\;op\;f_2 =
-\overline{x} \land \left(f_1|_{x\leftarrow 0}\;op\;f_2|_{x\leftarrow 0}\right)
-\lor x \land \left(f_1|_{x\leftarrow 1}\;op\;f_2|_{x\leftarrow 1}\right)
+f_1 op f_2 =
+\overline{x_i} \land \left(f_1|_{x_i\leftarrow 0}\;op\;f_2|_{x_i\leftarrow 0}\right)
+\lor x_i \land \left(f_1|_{x_i\leftarrow 1}\;op\;f_2|_{x_i\leftarrow 1}\right)
 $
@@
 $
-f_1|_{x=f_2}=
+f_1|_{x_i=f_2}=
 f_2 \land f_1|_{x_i=1} \lor \overline{f_2} \land f_1|_{x_i=0}
@@
 \noindent
+The ITE operator $ITE(x, y_1, y_2)$ chooses $y_1$ if $x$ evaluates to true and
+$y_2$ otherwise. Computing $ITE(x, y_1, y_2)$ is accomplished by utilizing the
+following recursive form:
+
 The time complexity then improves to be
-$O\left(|V_{f_1}|^2 \cdot |V_{f_2}|\right)$~\cite{BRYANT86}.
-~\cite[p. 686]{BRYANT86} lists Bryant's full pseudocode.
+$O\left(|V_{f_1}|^2 \cdot |V_{f_2}|\right)$~\cite{BRYANT86} when using the ITE
+operator.~\cite[p. 686]{BRYANT86} lists Bryant's full pseudocode.
 
@@
 
-\tbd
+Accoring to Brace et al.~\cite{BRACE90}, the ITE operator forms the core of an
+efficient OBDD package implementation. \tbr
 

### 2015-09-13 paper.tex
Fix wording

@@
 
-However, this would result in an algorithm having a time complexity of 
+This would result in an algorithm having a worst case time complexity of 
 $O\left(|V_{f_1}|^2 \cdot |V_{f_2}|^2\right)$. Therefore, Bryant~\cite{BRYANT86}
@@
 \noindent
-The ITE operator $ITE(x, y_1, y_2)$ chooses $y_1$ if $x$ evaluates to true and
-$y_2$ otherwise. Computing $ITE(x, y_1, y_2)$ is accomplished by utilizing the
-following recursive form:
+For a given node $x$, the ITE operator $ITE(x, hi(x), lo(x))$ chooses $hi(x)$ if
+$x$ evaluates to true and $lo(x)$ otherwise. Computing $ITE(x, hi(x), lo(x))$ is
+accomplished by utilizing the following recursive form:
+\begin{center}
+$
+ITE(x, hi(x), lo(x)) =
+(x, ITE(), ITE())
+$
+\end{center}
 

### 2015-09-13 paper.tex
Revise Compose operation (ITE operator)

@@
 \noindent
-For a given node $x$, the ITE operator $ITE(x, hi(x), lo(x))$ chooses $hi(x)$ if
-$x$ evaluates to true and $lo(x)$ otherwise. Computing $ITE(x, hi(x), lo(x))$ is
-accomplished by utilizing the following recursive form:
+For a given node $x$, the ITE operator for three nodes, \newline
+$ITE(x, hi(x), lo(x))$, chooses $hi(x)$ if $x$ evaluates to true and $lo(x)$
+otherwise. Computing $ITE(x, hi(x), lo(x))$ is accomplished by utilizing a
+recursive formulation~\cite{BRACE90}. To shorten the recursive formula, $l$ and
+$h$ are used to represent $lo(x)$ and $hi(x)$, respectively.
 \begin{center}
 $
-ITE(x, hi(x), lo(x)) =
-(x, ITE(), ITE())
+ITE(x, h, l) =
+(var(x), ITE(x|_{x=1}, h|_{x=1}, l|_{x=1}), ITE(x|_{x=0}, h|_{x=0}, l|_{x=0}))
 $
 \end{center}
+
+This formulation is based upon the fact, that every node $n$ of an OBDD represents
+a Boolean function and is denoted by a triple $(var(n), hi(n), lo(n))$. Brace et
+al.~\cite{BRACE90} called the variable associated with $n$, i.e. $x_n$ the
+\textit{top variable} of $n$.
 

### 2015-09-13 paper.tex
Fix consistent ITE formatting

@@
 already proposed to extend the \textit{Apply} operation to support ternary
-operations and then use the equivalent \textit{if-then-else} (\textit{ITE})
+operations and then use the equivalent \textit{if-then-else} ($ITE$)
 representation:
@@
 \noindent
-For a given node $x$, the ITE operator for three nodes, \newline
+For a given node $x$, the $ITE$ operator for three nodes, \newline
 $ITE(x, hi(x), lo(x))$, chooses $hi(x)$ if $x$ evaluates to true and $lo(x)$
@@
 The time complexity then improves to be
-$O\left(|V_{f_1}|^2 \cdot |V_{f_2}|\right)$~\cite{BRYANT86} when using the ITE
+$O\left(|V_{f_1}|^2 \cdot |V_{f_2}|\right)$~\cite{BRYANT86} when using the $ITE$
 operator.~\cite[p. 686]{BRYANT86} lists Bryant's full pseudocode.
@@
 implementing an OBDD framework efficiently, i.e. efficient data structures, the
-ITE operator, how to possibly approach the variable ordering problem and
+$ITE$ operator, how to possibly approach the variable ordering problem and
 \textit{garbage collection}. \tbr
@@
 As will be seen, \textit{Hash-based caches} are useful to provide an efficient
-implementation of the ITE operator. \tbr
+implementation of the $ITE$ operator. \tbr
 
@@
 
-Accoring to Brace et al.~\cite{BRACE90}, the ITE operator forms the core of an
+Accoring to Brace et al.~\cite{BRACE90}, the $ITE$ operator forms the core of an
 efficient OBDD package implementation. \tbr

### 2015-09-13 paper.tex
Add unique-table notion to hash table section
Finish ITE operator

@@
 
-This would result in an algorithm having a worst case time complexity of 
+This would result in an algorithm having a worst-case time complexity of 
 $O\left(|V_{f_1}|^2 \cdot |V_{f_2}|^2\right)$. Therefore, Bryant~\cite{BRYANT86}
@@
 each nonterminal node $n$ generated while executing the
-operation~\cite{BRYANT92}.
+operation~\cite{BRYANT92}. This is also referred to as
+\textit{unique-table}~\cite{BRACE90} imposes the canonical form of an OBDD. If a
+new node is generated, a corresponding entry is created in the unique-table.
 
@@
 Accoring to Brace et al.~\cite{BRACE90}, the $ITE$ operator forms the core of an
-efficient OBDD package implementation. \tbr
+efficient OBDD package implementation. Hence, they used a hash-based cache, called
+\textit{caching computed-table}, to implement it, exploiting a high locality of
+reference and consuming less memory.
+
+Compared to using a normal hash table, this results in more recursive calls to 
+$ITE$ (because of cache misses) but less time spent in garbage collection
+(described in Section~\ref{subsec:garbage-collection}) and look-ups (because no
+collision chains need to be searched).
+
+Nevertheless, it is worth noting that the worst-case time complexity increases
+to be exponential if all keys are hashed to the same value, compared to a
+polynomial time complexity for normal hash tables~\cite{BRACE90}.
+
+Furthermore, Brace et al.~\cite{BRACE90} advise to use \textit{standard triples},
+i.e. to define an $ITE(f_1, f_2, f_3)$-based equivalence relation on sets of
+three functions $f_1$, $f_2$ and $f_3$. This takes advantage of the fact that
+different parameters may yield the same result when $ITE$ is called. Hence, before
+the computed-table is accessed when executing $ITE$, the arguments are replaced
+by the standard arguments. Using this approach some recomputations can be avoided
+and, obviously, less space is consumed by the computed-table~\cite{BRACE90}.
+
+Brace et al.~\cite[p. 42]{BRACE90} provides the pseudocode of the $ITE$ algorithm
+as well as a modification, called $ite\_constant$, which can be used to implement
+logical implication tests more efficiently by avoiding the construction of
+intermediate nodes. \tbr
 

### 2015-09-13 paper.tex
Add Graph representation paragraph to Efficient data structures
Fix typo
Add complement edges to Graph representation

@@
 
+\paragraph*{Graph Representation}
+\mbox{} % workaround to get line break after paragraph
+
+To reduce the number of nodes of an OBDD, one can use \textit{complement edges}.
+A complement edge is an edge holding a single bit to indicate complementation.
+Hence, when $\overline{f}$ needs to be represented and a node $f$ is already
+present, a complement edge can be used to connect $f$ and thus avoid
+constructing the intermediate node $\overline{f}$. Also one of the two terminal
+nodes (e.g. \textbf{1}) can be removed by just using a complement edge to connect
+the other one (e.g. \textbf{0})~\cite{BRACE90}.
+
+Brace et al.~\cite{BRACE90} identified four functionally equivalent pairs of
+functions to be used for complement edges in order to maintain the canonical form.
+Let $x$ be the node having $lo(x)$ and $hi(x)$ on its outgoing 0- and 1-edge,
+respectively. Moreover, let $in(x)$ be the (one) incoming edge of node $x$ and
+$\overline{c}$ the complement of $c$. Then, the following table sums up the
+four pairs identified by Brace et al.~\cite{BRACE90}.
+
+\begin{center}
+    \begin{tabular}{lll}
+        $\left(in(x), lo(x), hi(x)\right)$ & $\Leftrightarrow$ & $\left(\overline{in(x)}, \overline{lo(x)}, \overline{hi(x)}\right)$ \tabularnewline
+        $\left(\overline{in(x)}, lo(x), hi(x)\right)$ & $\Leftrightarrow$ & $\left(in(x), \overline{lo(x)}, \overline{hi(x)}\right)$ \tabularnewline
+        $\left(in(x), \overline{lo(x)}, hi(x)\right)$ & $\Leftrightarrow$ & $\left(\overline{in(x)}, lo(x), \overline{hi(x)}\right)$ \tabularnewline
+        $\left(\overline{in(x)}, \overline{lo(x)}, hi(x)\right)$ & $\Leftrightarrow$ & $\left(in(x), lo(x), \overline{hi(x)}\right)$ \tabularnewline
+    \end{tabular}
+\end{center}
+
+Also, the 1-edge of each node must always be a regular one to guarantee a
+canonical form.
+
+Implementing complement edges enables one to compute the complement and identify
+complement functions in constant time (because $f$ and $\overline{f}$ use the same
+node). According to Brace et al.~\cite{BRACE90} this reduces the total runtime
+to form the OBBDD by almost a factor of 2 (for 12 examples). \tbr
+
 \subsection{ITE Operator}

### 2015-09-14 paper.tex
Add unique-table integration to graph representation
Start dynamic variable ordering

@@
 node). According to Brace et al.~\cite{BRACE90} this reduces the total runtime
-to form the OBBDD by almost a factor of 2 (for 12 examples). \tbr
+to form the OBBDD by almost a factor of 2 (for 12 examples).
+
+Another refinement to the graph representation, also proposed by Brace et
+al.~\cite{BRACE90}, is to integrate the unique-table into the graph representation,
+i.e. linking the collision chain of each node to the corresponding unique-table
+entry, to improve memory usage. \tbr
 
@@
 
+For an OBDD package it is desirable to provide the possibility to automatically 
+choose a good, though not optimal variable order for arbitrary functions. Of
+course, there exist heuristics to order variables before processing but
+Rudell~\cite{RUDELL93} proposed to integrate the determination and maintainance
+of the variable order into the OBDD package itself. This approach to the variable
+ordering problem is called \textit{dynamic variable ordering} and removes the
+necessity of choosing a variable ordering beforehand.
+
+Dynamic variable ordering executes the \textit{sifting} algorithm, also proposed
+by Rudell~\cite{RUDELL93}, to reorder the variables periodically and so minimize
+the size of the OBDD.
+
 \begin{figure}[ht]
@@
 \end{figure}
-\tbd
+\tbr
 

### 2015-09-14 paper.tex
Revise Dynamic variable ordering intro

@@
 of the variable order into the OBDD package itself. This approach to the variable
-ordering problem is called \textit{dynamic variable ordering} and removes the
-necessity of choosing a variable ordering beforehand.
+ordering problem is called \textit{dynamic variable ordering} and increases the
+robustness of an OBDD package by removing the necessity of choosing a proper
+variable ordering beforehand and thus running the risk of spacing out.
 
@@
 by Rudell~\cite{RUDELL93}, to reorder the variables periodically and so minimize
-the size of the OBDD.
+the size of the OBDD. This on-the-fly approach enabled Rudell~\cite{RUDELL93} to
+overcome hard problems for which heuristics did not find a variable ordering good
+enough to actually solve them, i.e. heuristics were able to solve 24 out of 35
+large circuits from the IWLS'91 and dynamic variable ordering using the sift
+algorithm was able to solve 33 out of these 35 large circuits~\cite{RUDELL93}.
+
+Periodically means, that a reordering is not performed by user requests but at
+appropriate points, determined by the OBDD package itself and so eliminate any
+user responsibility.
 

### 2015-09-14 paper.tex
Finish dynamic variable ordering basics

@@
 
+The core forms an efficient variable swap implementation, i.e. keep the changes
+in the OBDD as local as possible when swapping two adjacent variables $x$ and
+$y$. It has been found that swapping variable of two adjacent levels $i$ and
+$(i+1)$ only affects the nodes of these two levels. Hence, Rudell~\cite{RUDELL93}
+proposed to keep a separate unique-table for each level of the OBDD, similar to
+the last refinement in \textit{Graph representation} of
+Section~\ref{subsec:efficient-data-structures}. In essence, keeping separate 
+unique-tables for each node is the best solution.
+
+To ensure that each node represents the same function before and after the variable
+swap, Rudell~\cite{RUDELL93} proposed a node overwriting strategy. A given node
+$(var(x_i), hi(x_i), lo(x_i))$ at level $i$ is overwritten by the triple
+\begin{center}
+$(var(x_{i+1}), (x_i, h|_{x=1}, l|_{x=1}), (x_i, h|_{x=0}, l|_{x=0}))$
+\end{center}
+Here the notation of \textit{Compose} (Section~\ref{sec:operations}) is reused
+to simplify the formula, i.e. $h$ ($l$) denotes $hi(x_i)$ ($lo(x_i)$).
+
+Rudell~\cite{RUDELL93} also proposed to clear the computed-table and issue a
+garbage collection before minimizing the OBDD. Afterwards, all nodes of level
+$(i+1)$ can be freed incrementally if there exist no references from elsewhere than
+level $i$.
+
+Unfortunately, dynamic variable ordering has some bad consequences as well, namely
+\textit{negate-input-} and \textit{negate-else-edges} cannot be used without
+loosing the (desired) local complexity of variable swapping.
+
 \begin{figure}[ht]

### 2015-09-14 paper.tex
Add sifting algorithm intro

@@
 garbage collection before minimizing the OBDD. Afterwards, all nodes of level
-$(i+1)$ can be freed incrementally if there exist no references from elsewhere than
-level $i$.
+$(i+1)$ can be freed incrementally if there exist no references from elsewhere
+than level $i$.
+
+To minimize the OBDD, a novel minimization algorithm was introduced by
+Rudell~\cite{RUDELL93}, referred to as \textit{sifting algorithm}. It tries to
+find the ideal position of one variable when all other variables keep their
+positions. By definition, the ideal position of a variable is the position which
+minimizes the size of the OBDD.
 

### 2015-09-14 paper.tex
Add sifting algorithm description

@@
 
+The sifting algorithm therefore uses a brute-force enumeration strategy. It starts
+by swapping the processed variable $x$ downwards one level at a time, i.e. swapping 
+it with its successor variable, which is located on the layer below in the OBDD.
+This is done until $x$ is the next to last variable in the OBDD. Then, this is
+done in the other way round, i.e. swapping $x$ with its predecessor variable until
+$x$ becomes the top variable of the OBDD. While proceeding, the ideal position
+is memorized and after $x$ became the top variable, $x$ is moved to this ideal
+position by, again, swapping levels one by one.
+
 Unfortunately, dynamic variable ordering has some bad consequences as well, namely

### 2015-09-14 paper.tex
Add sifting algorithm example; Revise dynamic variable ordering description

@@
 \usepackage{subfigure}
+\usepackage{colortbl}
 
@@
 \newcommand{\tbd}{\textbf{[TO BE DONE]}}
+
+\newcommand\diag[4]{%
+  \multicolumn{1}{p{#2}|}{\hskip-\tabcolsep
+  $\vcenter{\begin{tikzpicture}[baseline=0,anchor=south west,inner sep=#1]
+  \path[use as bounding box] (0,0) rectangle (#2+2\tabcolsep,\baselineskip);
+  \node[minimum width={#2+2\tabcolsep-\pgflinewidth},
+        minimum  height=\baselineskip+\extrarowheight-\pgflinewidth] (box) {};
+  \draw[line cap=round] (box.north west) -- (box.south east);
+  \node[anchor=south west] at (box.south west) {#3};
+  \node[anchor=north east] at (box.north east) {#4};
+ \end{tikzpicture}}$\hskip-\tabcolsep}}
 
@@
 
-The sifting algorithm therefore uses a brute-force enumeration strategy. It starts
-by swapping the processed variable $x$ downwards one level at a time, i.e. swapping 
-it with its successor variable, which is located on the layer below in the OBDD.
-This is done until $x$ is the next to last variable in the OBDD. Then, this is
-done in the other way round, i.e. swapping $x$ with its predecessor variable until
-$x$ becomes the top variable of the OBDD. While proceeding, the ideal position
-is memorized and after $x$ became the top variable, $x$ is moved to this ideal
-position by, again, swapping levels one by one.
+The sifting algorithm therefore uses a brute-force enumeration strategy to find
+the optimal variable position. It starts by swapping the processed variable $x$
+downwards one level at a time, i.e. swapping it with its successor variable, which
+is located on the layer below in the OBDD. This is done until $x$ is the next to
+last variable in the OBDD. Then, this is done in the other way round, i.e. swapping
+$x$ with its predecessor variable until $x$ becomes the top variable of the OBDD.
+While proceeding, the ideal position is memorized and after $x$ became the top
+variable, $x$ is moved to this ideal position by, again, swapping levels one by
+one~\cite{RUDELL93}. Figure~\ref{fig:sifting-algorithm} shows this procedure
+applied to an example, requiring a total of 9 steps for 5 possible positions.
+
+\begin{figure}
+    \centering
+    \begin{tabular}{l|l|l|l|l|l|}
+        \hline
+        \diag{0.1em}{1cm}{\scriptsize{Step}}{\scriptsize{Pos}} & $1$ & $2$ & $3$ & $4$ & $5$ \tabularnewline
+        \hline
+        init & $v$ & $w$ & $x$ & $y$ & $z$ \tabularnewline
+        \hline
+        1 & $v$ & $w$ & \cellcolor{red!50}$y$ & \cellcolor{red!50}$x$ & $z$ \tabularnewline
+        \hline
+        2 & $v$ & $w$ & $y$ & \cellcolor{red!50}$z$ & \cellcolor{red!50}$x$ \tabularnewline
+        \hline
+        3 & $v$ & $w$ & $y$ & \cellcolor{red!50}$x$ & \cellcolor{red!50}$z$ \tabularnewline
+        \hline
+        4 & $v$ & $w$ & \cellcolor{red!50}$x$ & \cellcolor{red!50}$y$ & $z$ \tabularnewline
+        \hline
+        5 & $v$ & \cellcolor{red!50}$x$ & \cellcolor{red!50}$w$ & $y$ & $z$ \tabularnewline
+        \hline
+        6 & \cellcolor{red!50}$x$ & \cellcolor{red!50}$v$ & $w$ & $y$ & $z$ \tabularnewline
+        \hline\hline
+        7 & \cellcolor{green!50}$v$ & \cellcolor{green!50}$x$ & $w$ & $y$ & $z$ \tabularnewline
+        \hline
+        8 & $v$ & \cellcolor{green!50}$w$ & \cellcolor{green!50}$x$ & $y$ & $z$ \tabularnewline
+        \hline
+        9 & $v$ & $w$ & \cellcolor{green!50}$y$ & \cellcolor{green!50}$x$ & $z$ \tabularnewline
+        \hline
+    \end{tabular}
+    \caption{Sifting algorithm (ideal position of $x$: 4)}
+    \label{fig:sifting-algorithm}
+\end{figure}
+
+This procedure could be refined to choose the direction before actually start
+swapping the processed variable. As a result, the variable is always swapped
+towards the closer variable position first.
+
+To obtain a proper variable ordering, variables are sorted descending in the number
+of nodes per OBDD level. Then, the ideal position is computed for each variable
+utilizing the approach described above. The sifting algorithm has one big advantage
+over other similar algorithms, like \textit{window permutation algorithm}, namely
+variables can move many levels because all positions are evaluated, regardless of
+the temporary size of the OBDD.
 
@@
     \includestandalone[height = 3cm]{figs/min-obdd}
+    \caption{Minimized OBDD (optimal).}
     \label{fig:min-obdd}
-    \caption{Minimized OBDD (optimal).}
 \end{figure}

### 2015-09-14 paper.tex
Add complexity analysis of the dynamic variable ordering

@@
 variables can move many levels because all positions are evaluated, regardless of
-the temporary size of the OBDD.
+the temporary size of the OBDD~\cite{RUDELL93}.
+
+According to Rudell~\cite{RUDELL93}, $O(n^2)$ swaps of adjacent levels are
+necessary. Moreover, each swap has a complexity proportional to the with of the
+OBDD. Thus, Rudell~\cite{RUDELL93} advises to terminate whenever the OBDD doubles
+its size.
 

### 2015-09-14 paper.tex
Add pro and contra dynamic variable ordering

@@
 OBDD. Thus, Rudell~\cite{RUDELL93} advises to terminate whenever the OBDD doubles
-its size.
+its size. On the one hand, in contrast to heuristics and similar approaches, i.e.
+the window permutation algorithm, the dynamic variable ordering using the sifting
+algorithm has a significantly higher runtime. On the other hand, dynamic variable
+ordering using the sifting algorithm was able to complete operations which were
+not completed without it. Furthermore, the resulting OBDDs were significantly
+smaller~\cite{RUDELL93}.
 

### 2015-09-14 paper.tex
Add paragraph describing the minimized OBDD example

@@
 one~\cite{RUDELL93}. Figure~\ref{fig:sifting-algorithm} shows this procedure
-applied to an example, requiring a total of 9 steps for 5 possible positions.
+applied to an example, requiring a total of 9 steps for 5 possible positions. The
+first phase, determining the ideal position by swapping two variables at a time,
+is marked red, i.e. the two swapped variables are marked. The variable pairs
+swapped during the second phase, i.e. moving the $x$ to its ideal position (4),
+are marked green.
 
@@
 
+However, dynamic variable ordering is a good practice to obtain minimized OBDDs.
+For example, Figure~\ref{fig:min-obdd} shows the minimized, and thus optimal, OBDD
+representation of \newline $f=\left(A \land B\right) \lor \left(B \land C\right)$
+which, compared to the OBDD shown in Figure~\ref{subfig:obdd}, consists of less
+nodes.
+
 \begin{figure}[ht]

### 2015-09-14 paper.tex
Move Garbage collection section in front of Dynamic variable ordering
Add Garbage collection intro

@@
 
+\subsection{Garbage Collection}
+\label{subsec:garbage-collection}
+
+The first OBDD package containing kind of a garbage collection was the one
+introduced by Brace, Bryant and Rudell~\cite{BRACE90}. The garbage collection was
+based upon the reference counting technique, i.e. every node of the OBDD had a
+reference count. This reference count represents two kinds of references, (1)
+the number of other nodes and (2) the number of formulas referencing it, but does
+not take references from the unique- and computed-table into account. \tbr
+
 \subsection{Dynamic Variable Ordering}
@@
 
-\subsection{Garbage Collection}
-\label{subsec:garbage-collection}
-
-\tbd
-
 \section{Limitations}

### 2015-09-14 paper.tex
Add descriptions of reference count and dead nodes

@@
 the number of other nodes and (2) the number of formulas referencing it, but does
-not take references from the unique- and computed-table into account. \tbr
+not take references from the unique- and computed-table into account.
+
+Today, garbage collecion is a well-known topic. It is used to free up unnecessary
+allocated or unused memory. To accomplish this, the garbage collection has to know
+about which parts of the memory are \textit{dead} and thus can be freed. A node
+of an OBDD is called \textit{dead} if the reference count of it is equal to zero.
+The reference count of a node $x$ is decremented every time a corresponding
+formula is freed. If the reference count of $x$ reaches zero, the reference counts
+of $lo(x)$ and $hi(x)$ are decremented recursively.\tbr
 

### 2015-09-14 paper.tex
Add reclaim operation

@@
 formula is freed. If the reference count of $x$ reaches zero, the reference counts
-of $lo(x)$ and $hi(x)$ are decremented recursively.\tbr
+of $lo(x)$ and $hi(x)$ are decremented recursively.
+
+Instead of freeing dead nodes instantaneously, there exists a \textit{reclaim}
+operation which is performed whenever a lookup in the computed-table returns a
+dead node $x$. Such a reclaim operation increments the reference count of $x$ and
+propagates the reclaim operation to $lo(x)$ and $hi(x)$, respectively.\tbr
 

### 2015-09-14 paper.tex
Add citations

@@
 formula is freed. If the reference count of $x$ reaches zero, the reference counts
-of $lo(x)$ and $hi(x)$ are decremented recursively.
+of $lo(x)$ and $hi(x)$ are decremented recursively~\cite{BRACE90}.
 
@@
 dead node $x$. Such a reclaim operation increments the reference count of $x$ and
-propagates the reclaim operation to $lo(x)$ and $hi(x)$, respectively.\tbr
+propagates the reclaim operation to $lo(x)$ and $hi(x)$,
+respectively~\cite{BRACE90}. \tbr
 

### 2015-09-14 paper.tex
Add garbage collection invocation cases

@@
 propagates the reclaim operation to $lo(x)$ and $hi(x)$,
-respectively~\cite{BRACE90}. \tbr
+respectively~\cite{BRACE90}.
+
+The garbage collection is executed whenever the load factor (the ratio between
+number of keys stored and number of available bins) exceeds 4 after an insertion
+and $10\%$ of all OBDD nodes are dead. A garbage collection run deletes all entries
+in the computed-table referencing dead nodes, and frees all dead nodes. The sizes
+of computed- as well as unique-table are increased if the garbage collection does
+not find enough dead nodes. Another event invoking the garbage collection is the
+memory overflow of computed- and unique-table, although, the $10\%$-threshold of
+dead nodes is used again. In any other case, the garbage collection gives up (and
+of course, frees all intermediate nodes). \tbr
 

### 2015-09-14 paper.tex
Add effectiveness notice

@@
 memory overflow of computed- and unique-table, although, the $10\%$-threshold of
-dead nodes is used again. In any other case, the garbage collection gives up (and
-of course, frees all intermediate nodes). \tbr
+dead nodes is used again. In any other case, the garbage collection gives up (and,
+of course, frees all intermediate nodes)~\cite{BRACE90}.
+
+As stated by Brace et al.~\cite{BRACE90}, this provides effective memory usage
+at very low run-time costs, i.e. $3-7\%$ of the total runtime. \tbr
 

### 2015-09-14 paper.tex
Add limitations intro
Add integer multiplication part to Limitations

@@
 
-\tbd
+Some limitation were already mentioned briefly. This section provides a deeper
+insight on what cannot be accomplished using the standard OBDD representation.
+This includes classes of Boolean function which cannot be computed efficiently
+using OBDDs and, again, the variable ordering problem.
+
+Already in 1986, Bryant~\cite{BRYANT86} noticed that some classes of Boolean
+functions cannot be represented by an OBDD efficiently, regardless of the variable
+ordering chosen. He then studied the complexity of Boolean functions in
+1991~\cite{BRYANT91} to provide a lower bound on the OBDD as data structure (and
+very large scale integrated circuits). The complexity was expressed in the minimum
+size of the OBDD, i.e. the minimal number of nodes for any possible variable
+ordering $\pi$. Bryant~\cite{BRYANT91} mainly focused on the set of integer
+multipliers having word size $n$.
+
+Bryant~\cite{BRYANT91} first generalized a function over $2n$ variables. The size
+of an OBDD then can range from $2n$ up to $2^{n+1}-2$ for different variable
+orderings. Now, analyzing $n$-bit integer multiplication, Bryant~\cite{BRYANT86}
+already proved that for any variable ordering, there exists an $0 \le i \le 2n$
+such that the function representing output $i$ requires an OBDD of size
+$\Omega(1.09^n)$. Bryant himself~\cite{BRYANT91} showed that there is not much
+of a difference between computing the single output $(n-1)$ or the entire product
+of an $n$-bit integer multiplier. Bryant~\cite{BRYANT91} found the $n$-bit integer
+multiplication to be especially hard, requiring an OBDD having at least $2^{n/8}$
+nodes to compute the middle bit of the built product for any variable ordering.
+Hence, the complexity of an OBDD representing integer multiplication is
+exponential in $n$, where $n$ denotes the word size of the multiplied integers.
+
+\tbr
 

### 2015-09-14 paper.tex
Add new bound using universal hashing intro

@@
 
-\tbr
+Nevertheless, researches put some effort in improving this lower bound. In 2001,
+Woelfel~\cite{WOELFEL01} published a new bound lower bound as well as a first 
+nontrivial upper bound for the size of OBDDs representing $n$-bit integer
+multiplication by utilizing the concept of \textit{universal hashing}. \tbr
 

### 2015-09-14 paper.tex
Add full proof reference to BRYANT91
Finish integer multiplications part of Limitations

@@
 exponential in $n$, where $n$ denotes the word size of the multiplied integers.
+One can read the full proof in~\cite{BRYANT91}.
 
@@
 nontrivial upper bound for the size of OBDDs representing $n$-bit integer
-multiplication by utilizing the concept of \textit{universal hashing}. \tbr
+multiplication by utilizing the concept of \textit{universal hashing}. Therefore,
+he utilized a universal family of hash functions to improve the exponential lower
+bound on the OBDD-size of integer multiplication to
+$2^{\lfloor\frac{n}{2}\rfloor}/96$. Moreover, Woelfel~\cite{WOELFEL01} proved
+the upper bound for the OBDD-size of computing the middle bit of an $n$-bit
+multiplication to be $7/3 \cdot 2^{\frac{4 \cdot n}{3}}$.
+
+Nevertheless, all lower bounds for $n$-bit integer multiplication are exponential
+in the number of nodes. Especially because of the high practical relevance of
+$n$-bit integer multiplication this is one of the main drawbacks of the OBDD
+representation~\cite{BRYANT86, BRYANT91, WOELFEL01}. \tbr
 

### 2015-09-14 paper.tex
Add NP-completeness notice of improving the variable ordering

@@
 $n$-bit integer multiplication this is one of the main drawbacks of the OBDD
-representation~\cite{BRYANT86, BRYANT91, WOELFEL01}. \tbr
+representation~\cite{BRYANT86, BRYANT91, WOELFEL01}.
+
+Another topic researchers put much effort into was the variable ordering problem.
+Section~\ref{subsec:dynamic-variable-ordering} already described one possible
+approach to this problem but there are many other approaches such as heuristics
+and simulated annealing algorithms~\cite{BOLLIG96}.
+
+Bollig and Wegener~\cite{BOLLIG96} studied the complexity of improving the
+variable ordering. They investigated this problem by analyzing the following
+problem: Given an OBDD representing $f$ and a size bound $s$. Does an OBDD exist,
+also representing $f$ and respecting an arbitrary variable ordering, having at
+most $s$ nodes? "Unfortunately", they were able to prove this problem to be 
+NP-complete. Hence, improving the variable ordering of OBDDs is
+NP-complete~\cite{BOLLIG96}. \tbr
 

### 2015-09-14 paper.tex
Add section Symbolic model checking using OBDDs
Fix heading
Add Symbolic model checking intro

@@
 
+\section{Symbolic Model Checking}
+\label{sec:symbolic-model-checking}
+
+One specific application of reduced ordered binary decision diagrams is 
+\textit{symbolic model checking}, introduced by Burch et al.~\cite{BURCH90} in
+1990. \textit{Symbolic}, because the state space is not represented explicitly
+but symbolically by utilizing OBDDs to represent $\mu$-Calculus formulas.
+In~\cite{BURCH90}, Burch, Clarke, McMillan, Dill and Hwang derived an efficient
+decision procedures for CTL model checking as well as satisfiability of LTL
+formulas, strong and weak observational equivalence of finite transition systems
+and language containment for finite $\omega$-automata by introducing a new
+$\mu$-Calculus model checking algorithm. However, this survey just outlines the
+$\mu$-Calculus and the efficient decision procedures for CTL model checking.
+
+The main idea behind this approach is to represent the state space symbolically,
+instead of explicitly with e.g. a state table. Such explicit state tables may grow
+exponentially with the number of concurrent components, because the state space
+does so. Commonly, this is referred to as the
+\textit{state explosion problem}~\cite{BURCH90, MCMILLAN92}. \tbr
+
 \section{Alternative Representations}

### 2015-09-14 paper.tex
Add OBDD representation for Symbolic model checking

@@
 does so. Commonly, this is referred to as the
-\textit{state explosion problem}~\cite{BURCH90, MCMILLAN92}. \tbr
+\textit{state explosion problem}~\cite{BURCH90, MCMILLAN92}. To overcome this
+problem, Burch et al.~\cite{BURCH90} as well as McMillan~\cite{MCMILLAN92}
+proposed to utilize reduced ordered binary decision diagrams to represent the 
+state space symbolically. Although OBDDs cannot avoid a state explosion in general,
+many practically relevant systemes with large state spaces can be
+verified~\cite{BURCH90, MCMILLAN92}. Experiments~\cite{BURCH90} showed that
+symbolic model checking enables to verify models with more than $10^{20}$ states,
+whereas explicit enumeration approaches can only verify models up to $10^6$
+states.
+
+Burch et al.~\cite{BURCH90} and McMillan~\cite{MCMILLAN92} extend Bryant's OBDD
+representation~\cite{BRYANT86} by quantification over Boolean variables (QBF) and
+substitution of variable names. The quantification over Boolean variables is
+realized by using Bryant's \textit{Restrict} operation (see
+Section~\ref{sec:operations}). Let $f$ be a QBF formula and $x$ be a Boolean
+variable. The QBF formula $\exists x\left(f\right)$ then is equivalent to
+$f|_{a=0} \lor f|_{a=1}$~\cite{BURCH90}. Substitution of variables, e.g.
+substitute $x$ by $y$ (denoted $f\langle x\leftarrow y\rangle$), is computed using
+the following equivalence:
+\begin{center}
+$
+f\langle x\leftarrow y\rangle \equiv
+\exists x\lbrack\left(x \Leftrightarrow y\right) \land f\rbrack
+$
+\end{center}
+\tbr
 

### 2015-09-15 paper.tex
Fix wording
Add mu-Calculus intro and definitions

@@
 $f|_{a=0} \lor f|_{a=1}$~\cite{BURCH90}. Substitution of variables, e.g.
-substitute $x$ by $y$ (denoted $f\langle x\leftarrow y\rangle$), is computed using
-the following equivalence:
+substitution of $x$ by $y$ (denoted $f\langle x\leftarrow y\rangle$), is computed
+using the following equivalence:
 \begin{center}
@@
 \end{center}
+
+Before discussing the symbolic model checking algorithms \texttt{BDD}$_f$ and
+\texttt{BDD}$_R$, the used $\mu$-Calculus has to be described. For a given finite
+signature $\mathcal{S}$, each symbol is either an individual or a predicate
+variable having positive arity. Burch et al.~\cite{BURCH90} considered two
+syntactic categories, i.e. \textit{formulas} and \textit{relational terms}, having
+the following forms.
+
+\noindent
+\textbf{Formulas:}~\cite{BURCH90}
+\begin{enumerate}
+    \item{\textbf{True}, \textbf{False}}.
+    \item{
+        $\lbrack z_1 = z_2 \rbrack$; $z_{1, 2}\in\mathcal{S}$ are individual
+        variables.
+    }
+    \item{
+        $\neg f$, $f \lor g$, $\exists z\left( f \right)$; $z\in\mathcal{S}$ is
+        an individual variable, $f$ and $g$ are formulas.
+    }
+    \item{
+        $P\left(z_1, \ldots, z_n\right)$; $P$ is an $n$-ary relational term and
+        \newline $z_{1, \ldots, n}\in\mathcal{S}$ are individual variables not
+        free in $P$.
+    }
+\end{enumerate}
+\textbf{$n$-ary relational terms:}~\cite{BURCH90}
+\begin{enumerate}
+    \item{$Z$; $Z\in\mathcal{S}$ is an $n$-ary predicate variable.}
+    \item{
+        $\lambda z_1, \ldots, z_n\left( f \right)$; $f$ is a formula and
+        $z_1, \ldots, z_n\in\mathcal{S}$ are distinct individual variables.
+    }
+    \item{
+        $\mu Z \left( P \right)$; $Z\in\mathcal{S}$ is an $n$-ary predicate variable
+        and $P$ is an $n$-ary relational term \textit{formally monotone in} $Z$,
+        i.e. all free occurrences of $Z$ in $P$ have to be evenly negated. This
+        $n$-ary relatioinal term is the \textit{least fixed point} (\textit{lfp})
+        of an $n$-ary relational term $P$. Analogous, $\nu Z \left( P \right)$ is
+        the \textit{greatest fixed point} (\textit{gfp}) and is equivalent to
+        $\neg\mu Z \left( \neg P \langle Z \leftarrow \left( \neg Z \right) \rangle \right)$.
+    }
+\end{enumerate}
+
 \tbr

### 2015-09-15 paper.tex
Add semantic function D and its mappings

@@
 
+To determine whether a formula is \textbf{True} or \textbf{False}, Burch et
+al.~\cite{BURCH90} defined a structure $\mathcal{M}\left( D, I_P, I_D \right)$,
+where $D$, $I_P$ and $I_D$ represent the (non-empty) \textit{domain} of
+$\mathcal{M}$, the \textit{relational variable interpretation} and the
+\textit{individual variable interpretation}, respectively. Furthermore, a semantic
+function $\mathcal{D}$ is introduced, mapping formulas to elements of
+\begin{center}
+$\left( I_P \rightarrow \left( I_D \rightarrow \lbrace true, false \rbrace \right) \right)$
+\end{center}
+and $n$-ary relational terms to elements of
+\begin{center}
+$\left( I_P \rightarrow \left( I_D \rightarrow 2^{D^n} \right) \right)$
+\end{center}.
+
 \tbr

### 2015-09-15 paper.tex
Add inductive definition of semantic function D

@@
 
+Burch et al.~\cite{BURCH90} then defined $\mathcal{D}$ inductively on formulas
+and relational terms. For a structure $\mathcal{M}$ and a formula $f$,
+$\mathcal{M}$ satisfies $f$, written as $\mathcal{M} \models f$, according to
+these semantics:
+\begin{enumerate}
+    \item{
+        $\mathcal{D} \left( z_1 = z_2 \right) \left( I_P \right) \left( I_D \right) :=
+        \left( I_D \left( z_1 \right) = I_D \left( z_2 \right) \right)$
+    }
+    \item{
+        $\mathcal{D} \left( \neg f_1 \right) \left( I_P \right) \left( I_D \right) :=
+        \neg \left( \mathcal{D} \left( f_1 \right) \left( I_P \right) \left( I_D \right) \right)$
+    }
+    \item{
+        $\mathcal{D} \left( f_1 \lor f_2 \right) \left( I_P \right) \left( I_D \right) := \newline
+        \mathcal{D} \left( f_1 \right) \left( I_P \right) \left( I_D \right) \lor\;\mathcal{D} \left( f_2 \right) \left( I_P \right) \left( I_D \right)$
+    }
+    \item{
+        $\mathcal{D} \left( \exists z \left( f_1 \right) \right) \left( I_P \right) \left( I_D \right) := \newline
+        \exists e \in D . \left( \mathcal{D} \left( f_1 \right) \left( I_P \right) \left( I_D \langle z \leftarrow e \rangle \right) \right)$
+    }
+    \item{
+        $\mathcal{D} \left( P\left( z_1, \ldots, z_n \right) \right) := \newline
+        \langle I_D \left( z_1 \right), \ldots, I_D \left( z_n \right) \rangle \in \mathcal{D} \left( P \right) \left( I_P \right) \left( I_D \right)$
+    }
+    \item{
+        $\mathcal{D} \left( Z \right) \left( I_P \right) \left( I_D \right) :=
+        I_P \left( Z \right)$
+    }
+    \item{
+        $\mathcal{D} \left( \lambda z_1, \ldots, z_n \left( f \right) \right) \left( I_P \right) \left( I_D \right) := \newline
+        \lbrace \langle e_1, \ldots, e_n \rangle : \mathcal{D} \left( f \right) \left( I_P \right) \left( I_D \langle z_1 \leftarrow e_1, \ldots, z_n \leftarrow e_n \rangle \right) \rbrace$
+    }
+    \item{
+        $\mathcal{D} \left( \mu Z \left( P \right) \right) :=
+        \text{lfp}\;\lambda Q \in 2^{D^n} . \left( \mathcal{D} \left( P \right) \left( I_P \langle Z \leftarrow Q \rangle \right) \left( I_D \right) \right)$,
+        where lfp is the least fixed point w.r.t. the inclusion ordering.
+    }
+\end{enumerate}
+
  \tbr

### 2015-09-15 paper.tex
Add model checking algorithms intro

@@
 
- \tbr
+Finally, after defining the $\mu$-Calculus formally, the model checking algorithms
+can be described. They operate on the limited Boolean domain, i.e.
+$D = \lbrace 0, 1 \rbrace$ but Burch et al.~\cite{BURCH90} also described how any
+\textit{finite} domain can be encoded into this limited Boolean domain. This last
+fact generalizes the introduced algorithms to any \textit{finite} domain. \tbr
 

### 2015-09-15 paper.tex
Add BDDf and OBDD relationship

@@
 
-Before discussing the symbolic model checking algorithms \texttt{BDD}$_f$ and
-\texttt{BDD}$_R$, the used $\mu$-Calculus has to be described. For a given finite
+Before discussing the symbolic model checking algorithms $\text{BDD}_f$ and
+$\text{BDD}_R$, the used $\mu$-Calculus has to be described. For a given finite
 signature $\mathcal{S}$, each symbol is either an individual or a predicate
@@
 \textit{finite} domain can be encoded into this limited Boolean domain. This last
-fact generalizes the introduced algorithms to any \textit{finite} domain. \tbr
+fact generalizes the introduced algorithms to any \textit{finite} domain.
+
+They described two function representing the algorithm:
+$\text{BDD}_f \left( f, I_P \right)$ and $\text{BDD}_R \left( R, I_P \right)$,
+where $f$ and $R$ denote a formula and a relational term, respectively. For 
+$\text{BDD}_f$, $I_P$ represents a relational variable interpretation assigning
+values to free relational variables in $f$~\cite{BURCH90}.
+
+A given individual variable interpretation $I_D$ satisfies \newline
+$\text{BDD}_f \left( f, I_P \right)$ iff $f$ is satisfied by $\mathcal{M}$ having
+interpretations $I_P$ and $I_D$. OBDDs in combination with place-holder variables
+$d_1, \ldots, d_n$ are used to represent the values of the relational variables
+$\in I_P$. $d_i$ represents the $i^{th}$ argument of a relation and an $n$-ary
+relation holds for arguments $y_1, \ldots, y_n$ iff the interpretation
+$\langle d_1 \leftarrow y_1, \ldots, d_n \leftarrow y_n \rangle$ satisfies the
+OBDD.
+
+\tbr
 

### 2015-09-15 paper.tex
Add BDDf function description

@@
 
+The complete code of $\text{BDD}_f$ and $\text{BDD}_R$ is listed
+in~\cite[p. 432]{BURCH90}. However, both functions are described briefly in the
+following paragraphs.
+
+In essence, $\text{BDD}_f \left( f, I_P \right)$ distinguishes between five cases
+derived from the OBDD semantics and the $\mu$-Calculus, utilizing the graph-based
+algorithms of Bryant~\cite{BRYANT86} (also see Section~\ref{sec:operations}) as
+well as the $\text{BDD}_R$ function (for applying a relational term $R$) with a
+subsequent substitution of the place-holders with the argument variables, i.e.
+$\langle d_1 \leftarrow x_1, \ldots, d_n \leftarrow x_n \rangle$~\cite{BURCH90}.
+
 \tbr

### 2015-09-15 paper.tex
Add BDDR function description

@@
 
-The complete code of $\text{BDD}_f$ and $\text{BDD}_R$ is listed
-in~\cite[p. 432]{BURCH90}. However, both functions are described briefly in the
-following paragraphs.
-
 In essence, $\text{BDD}_f \left( f, I_P \right)$ distinguishes between five cases
@@
 $\langle d_1 \leftarrow x_1, \ldots, d_n \leftarrow x_n \rangle$~\cite{BURCH90}.
+
+The OBDD returned by $\text{BDD}_R \left( R, I_P \right)$ may contain both types
+of variables, the place-holder and the individual variables, because a relational
+term may contain free individual variables. Hence,
+$\text{BDD}_R \left( R, I_P \right)$ is satisfied iff
+$\mathcal{D} \left( R \right) \left( I_P \right) \left( I_D \right)$ holds for
+the $n$-tuple $\langle I_A \left( d_1 \right), \ldots, I_A \left( d_n \right) \rangle$
+for given $I_D$ and $I_A$, representing the interpretations for individual and
+place-holder variables, respectively.
+
+$\text{BDD}_R \left( R, I_P \right)$ distinguishes three cases:
+\begin{enumerate}
+    \item{If $R$ is a relational variable, $I_P \left( R \right)$ is returned.}
+    \item{
+        If $R$ is a $\lambda$-abstraction, an OBDD with substituted variables is
+        returned, i.e. $\langle x_1 \leftarrow d_1, \ldots, x_n \leftarrow d_n \rangle$.
+    }
+    \item{
+        If $R$ is fixed point computation, a helper function is called. This
+        helper function computes the fixed point by a series of approximations
+        $T_1, \ldots, T_k$, starting from an empty relation and terminating as
+        soon as $T_{i+1} = T_i$, which can be tested easily using OBDD because
+        equivalence testing can be done in constant time.
+    }
+\end{enumerate}
+
+The complete code of $\text{BDD}_f$ and $\text{BDD}_R$ is listed
+in~\cite[p. 432]{BURCH90}.
 

### 2015-09-15 paper.tex
Add CTL and Kripke structure descriptions

@@
 
+Based upon this two functions and the $\mu$-Calculus, one can now do CTL model
+checking by representing a CTL formula $f$ as a $\mu$-Calculus relational term
+$R$. Assuming standard CTL syntax and semantics, i.e. for the set of atomic
+propositions $AP$
+\begin{enumerate}
+    \item{$p \in AP$ is a CTL formula.}
+    \item{
+        Given two CTL formulas $f_1$ and $f_2$, also $\neg f_1$, $f_1 \land f_2$,
+        $\text{\textbf{EX}} f$,
+        $\text{\textbf{E}} \lbrack f_1 \text{\textbf{U}} f_2 \rbrack$ and
+        $\text{\textbf{EG}} f_1$ are CTL formulas.
+    }
+\end{enumerate}
+
+A CTL formula is represented by a \textit{Kripke structure} \newline
+$\mathcal{M} \left( \mathcal{S}, \mathcal{S}_0, R, L \right)$, where $\mathcal{S}$
+is a finite set of states over $AP$, $\mathcal{S}_0 \in \mathcal{S}$ is the set
+of initial states, $R \subseteq \mathcal{S} \times \mathcal{S}$ is the transition
+relation and $L: \mathcal{S} \rightarrow 2^{AP}$ is a labelling function. An
+infinite sequence of states $\pi = s_0s_1s_2\ldots$ is called \textit{path} if
+$\forall i \in \mathbb{N}.\left( s_i, s_{i+1} \right) \in R$.
+
 \tbr

### 2015-09-15 paper.tex
Restructure

@@
 The complete code of $\text{BDD}_f$ and $\text{BDD}_R$ is listed
-in~\cite[p. 432]{BURCH90}.
-
-Based upon this two functions and the $\mu$-Calculus, one can now do CTL model
+in~\cite[p. 432]{BURCH90}. Based upon these two functions and the $\mu$-Calculus, one can now do CTL model
 checking by representing a CTL formula $f$ as a $\mu$-Calculus relational term
-$R$. Assuming standard CTL syntax and semantics, i.e. for the set of atomic
+$R$.
+
+Assuming standard CTL syntax and semantics, i.e. for the set of atomic
 propositions $AP$

### 2015-09-15 paper.tex
Add EGf representation

@@
         Given two CTL formulas $f_1$ and $f_2$, also $\neg f_1$, $f_1 \land f_2$,
-        $\text{\textbf{EX}} f$,
+        $\text{\textbf{EX}} f_1$,
         $\text{\textbf{E}} \lbrack f_1 \text{\textbf{U}} f_2 \rbrack$ and
@@
 
+Now, equivalent $\mu$-Calculus relational terms need to be formulated for
+$\text{\textbf{EX}} f$,
+$\text{\textbf{E}} \lbrack f_1 \text{\textbf{U}} f_2 \rbrack$ and
+$\text{\textbf{EG}} f$. A CTL formula $f$ is \textbf{True} at state $s$ iff
+$R \left( s \right)$ is \textbf{True}.
+
+$\text{\textbf{EX}} f$ is satisfied for a state $s$ iff a state satisfying $f$
+exists and $R \left( s, t \right)$ holds. In terms of $\mu$-Calculus relational
+terms, $\text{\textbf{EX}} f$ can be formulated as
+$\lambda u \lbrack \exists v \left( f(v) \land R \left( u, v \right) \right) \rbrack$.
+
 \tbr

### 2015-09-15 paper.tex
Add EfUg description

@@
 terms, $\text{\textbf{EX}} f$ can be formulated as
-$\lambda u \lbrack \exists v \left( f(v) \land R \left( u, v \right) \right) \rbrack$.
-
-\tbr
+$\lambda u \lbrack \exists v \left( f(v) \land R \left( u, v \right) \right) \rbrack$~\cite{BURCH90}.
+
+The other two formulas are represented by fixed point characterizations, computed
+using direct iteration or iterative squaring (see~\cite[p. 433]{BURCH90}).
+
+$\text{\textbf{E}} \lbrack f_1 \text{\textbf{U}} f_2 \rbrack$ is satisfied for a
+state $s$ if $f_2$ is satisfied by $s$ or $s$ satisfies $f_1$ and $s$ has a
+successor state $t$ such that a path exists starting at $t$ and satisfying
+$f_1 \text{\textbf{U}} f_2$. The $\mu$-Calculus representation is defined by
+$\mu Q \lbrack f_2 \lor \left( f_1 \land \text{\textbf{EX}} Q \right) \rbrack$~\cite{BURCH90}.\tbr
 

### 2015-09-15 paper.tex
Add EGf description

@@
 successor state $t$ such that a path exists starting at $t$ and satisfying
-$f_1 \text{\textbf{U}} f_2$. The $\mu$-Calculus representation is defined by
-$\mu Q \lbrack f_2 \lor \left( f_1 \land \text{\textbf{EX}} Q \right) \rbrack$~\cite{BURCH90}.\tbr
+$f_1 \text{\textbf{U}} f_2$. The $\mu$-Calculus representation (lfp) is defined by
+$\mu Q \lbrack f_2 \lor \left( f_1 \land \text{\textbf{EX}} Q \right) \rbrack$~\cite{BURCH90}.
+
+$\text{\textbf{EG}} f$ is satisfied for a state $s$ if $s$ satisfies $f$ and there
+exists a successor state $t$ satisfying $f$ globally. Hence, the equivalent 
+gfp representation is
+$\nu Q \lbrack f \land \text{\textbf{EX}} Q \rbrack$~\cite{BURCH90}.
+
+\tbr
 

### 2015-09-15 paper.tex
Finish Symbolic model checking
Add Alternative representations intro

@@
 
-\tbr
+Utilizing these equivalent representations, one efficient CTL model checking.
+Moreover, Burch, Clarke, McMillan and Dill~\cite{BURCH90-2} described how to use
+these methods to do \textit{fair} CTL model checking. \tbr
 
@@
 
-\tbd
+In this last section, some alternative representations of BDDs are discussed
+briefly, e.g. BMDs or BEDs. These representations were introduced primarily
+because the OBDD representation is insufficient for some problems of practical
+relevance. Bryant~\cite{BRYANT95} already reviewed most alternative represenations
+but some were introduced after this overview, e.g. the
+\textit{Binary Expression Diagrams} introduced by Andersen and
+Hulgaard~\cite{ANDERSEN97}. \tbr
 

### 2015-09-15 paper.tex
Fix wording
Add Alternative representations paragraphs
Add FDD description

@@
 
-In 1986 Bryant\cite{BRYANT86} published a paper of high impact, describing a
+In 1986 Bryant\cite{BRYANT86} published a paper of strong impact, describing a
 \textit{reduced} and \textit{ordered} class of BDDs, so-called \textit{ROBDDs}
@@
 
+\paragraph*{Free BDDs}
+\mbox{} % workaround to get line break after paragraph
+
+Although the variable ordering problem is well-studied, OBDDs have some severe
+limitations for some classes of Boolean functions, regardless of the variable
+ordering chosen, e.g. integer multiplication (see Section~\ref{sec:limitations}).
+One way researchers tried to overcome this issue is relaxing the variable ordering
+requirement. \textit{Free BDDs} (\textit{FDDs}) is one proposed approach. FDDs
+do not require the variables in the BDD to be ordered but, instead, require that
+each variable only occurs once along any path (starting from the root node).
+Hence, FDDS are also referred to as \textit{1-time branching programs}. Moreover,
+FDDs have the same desirable algorithmic properties as OBDDs~\cite{BRYANT95}. \tbr
+
+\paragraph*{Zero-Suppressed BDDs}
+\mbox{} % workaround to get line break after paragraph
+
+\tbd
+
+\paragraph*{Binary Moment Diagrams}
+\mbox{} % workaround to get line break after paragraph
+
+\tbd
+
+\paragraph*{Binary Expression Diagrams}
+\mbox{} % workaround to get line break after paragraph
+
+\tbd
+
 % balance columns on last page

### 2015-09-15 paper.tex
Add ZBDD description

@@
 
-\tbd
+This representation is especially useful when the data can be represented as
+(sparse) bit vectors. Any subset of these bit vectors can be represented by a
+Boolean function over $n$ variables. If the corresponding bit vector is in the
+set, the \textit{Zero-Suppressed BDDs} (\textit{ZBDDs}) yields 1 for a given
+variable assignment.
+
+The bit vector are \textit{sparse} in the sense that the set contains much less
+than $2^n$ vectors and the bit vectors themselves have many zero elements. ZBDDs
+are similar to OBDDs but they use a different reduction rule, exploiting the
+sparse bit vectors. In ZBDDs, nodes can be eliminated whenever setting the
+corresponding variable to 1 yields 0 as result, which is often the case for sparse
+bit vectors. This representation proved successful to solve some combinatorial
+problems~\cite{BRYANT95}. \tbr
 

### 2015-09-15 paper.tex
Remove BMD paragraph
Add BED intro

@@
 bit vectors. This representation proved successful to solve some combinatorial
-problems~\cite{BRYANT95}. \tbr
-
-\paragraph*{Binary Moment Diagrams}
-\mbox{} % workaround to get line break after paragraph
-
-\tbd
+problems~\cite{BRYANT95}. \tbr \newpage
 
@@
 
-\tbd
+In 1997, Andersen and Hulgaard~\cite{ANDERSEN97} proposeded extending OBDDs with
+a third type of nodes representing \textit{operators}. The resulting
+representation is referred to as \textit{Binary Expression Diagram}
+(\textit{BED}). Such an operator node $n$ has one attribute $op(n)$, the binary
+Boolean operator, and, like nonterminal nodes, two children $lo(n)$ and $hi(n)$.
+Operator nodes represent their corresponding Boolean connectives in the BED graph.
+ \tbr
 

### 2015-09-15 paper.tex
Add BED operator node description

@@
 Operator nodes represent their corresponding Boolean connectives in the BED graph.
+
+Just like OBDDs, BEDs share equivalent Boolean subexpressions. Furthermore, BEDs
+have their own set of manipulating operations, especially because of the operator
+nodes. For example, the \textit{up-step} operation moves variables up in the BED
+hierarchy, if possible. As an effect, operator nodes are pushed towards the
+terminal nodes and some operator nodes may be evaluated if some of the associated
+expressions are terminal nodes. The BED may be simplified afterwards. These steps
+can be repeated to reduce the BED over and over again, and simplifying some tests,
+e.g. tautology testing. It is also possible to move all variable up in a single
+step. Nevertheless, reducing the BED step-by-step can exploit structural
+information.
+
 \tbr

### 2015-09-15 paper.tex
Add succinctness notice

@@
 Boolean operator, and, like nonterminal nodes, two children $lo(n)$ and $hi(n)$.
-Operator nodes represent their corresponding Boolean connectives in the BED graph.
+Operator nodes represent their corresponding Boolean connectives in the BED
+graph~\cite{ANDERSEN97}.
 
@@
 step. Nevertheless, reducing the BED step-by-step can exploit structural
-information.
+information~\cite{ANDERSEN97}.
+
+In contrast to OBDDs, BEDs are \textit{exponentially more succinct} because
+arbitrary circuits can be translated into BEDs. As a consequence, the BED
+representation suffices to solve integer multiplication using a quadratic number
+of nodes~\cite{ANDERSEN97}.
 

### 2015-09-15 paper.tex
Add BED reducedness notice

@@
 
+Analogous to OBDDs and FDDs, OBEDs and FBEDs can be defined, whereas OBEDs are
+OBDDs without operator nodes. The reduction rule of BEDs is similar to the one
+of OBDDs but extended with an additional rule for operator nodes. For any operator
+nodes $n$ it has to hold that $lo(n)$ and $hi(n)$ are nonterminals in order to
+guarantee reducedness. However, even if a BED is reduced, canonicity is not
+guaranteed (in contrast to OBDDs).
+
 \tbr

### 2015-09-15 paper.tex
Finish BEDs

@@
 nodes $n$ it has to hold that $lo(n)$ and $hi(n)$ are nonterminals in order to
-guarantee reducedness. However, even if a BED is reduced, canonicity is not
-guaranteed (in contrast to OBDDs).
-
-\tbr
-
+guarantee reducedness. Apart from that, before generating a new node, an existence
+check is performed. If there exists an equivalent node, this node is reused.
+Otherwise the new node is created and considered in the next check. However, even
+if a BED is reduced, canonicity is not guaranteed (in contrast to
+OBDDs)~\cite{ANDERSEN97}.
+
+Andersen and Hulgaard~\cite{ANDERSEN97} also introduced some checks especially
+for operator nodes in order to reduce the BED-size even further. Moreover a
+detailed description of how to convert a BED into an equivalent OBDD is provided,
+distinguishing between step-by-step variable moving (denoted as \textit{up\_one})
+and block-wise variable moving (denoted as \textit{up\_all}).
+
+The \textit{up\_one} operation can be implemented with expected runtime of $O(n)$,
+where $n$ denotes the number reachable of nodes from a given node. Nonetheless,
+because BEDs are exponentially more succinct than OBDDs, the complete
+transformation has an exponential worst-case complexity. Also \textit{up\_all} has
+an exponential worst-case complexity (for the same reason)~\cite{ANDERSEN97}.
+
+Andersen and Hulgaard~\cite{ANDERSEN97} concluded that \textit{up\_all} is a
+memory efficient approach to construct OBDDs in contrast to the OBDD construction
+from scratch, without actually increasing the runtime. \textit{up\_one} is used
+to implement existential quantification and substitution efficiently, having
+linear runtime complexity in the size of the BED. Reconsidering symbolic model
+checking, BEDs can be used to evaluate fixed point computations directly without
+converting it to an OBDD. Moreover, BEDs can perform fixed point computations more
+efficiently because of their succinct representation~\cite{ANDERSEN97}. \tbr
+
+\clearpage
 % balance columns on last page
-\balance
+%\balance
 

### 2015-09-15 paper.tex
Revise Shannon's expansion
Revise Compose formulas
Fix wording
Fix formatting

@@
 f_1 op f_2 =
-\overline{x_i} \land \left(f_1|_{x_i\leftarrow 0}\;op\;f_2|_{x_i\leftarrow 0}\right)
-\lor x_i \land \left(f_1|_{x_i\leftarrow 1}\;op\;f_2|_{x_i\leftarrow 1}\right)
+\overline{x_i} \cdot \left(f_1|_{x_i\leftarrow 0}\;op\;f_2|_{x_i\leftarrow 0}\right)
++ x_i \cdot \left(f_1|_{x_i\leftarrow 1}\;op\;f_2|_{x_i\leftarrow 1}\right)
 $
@@
 f_1|_{x_i=f_2}=
-f_2 \land f_1|_{x_i=1} \lor \overline{f_2} \land f_1|_{x_i=0}
+f_2 \cdot f_1|_{x_i=1} + \overline{f_2} \cdot f_1|_{x_i=0}
 $
@@
 $
-ITE\left(f_2, f_1|_{x_i=1}, f_1|_{x_i=0}\right) \equiv
-f_2 \land f_1|_{x_i=1} \lor \overline{f_2} \land f_1|_{x_i=0}
+ITE\left(f_2, f_1|_{x_i=1}, f_1|_{x_i=0}\right) =
+f_2 \cdot f_1|_{x_i=1} + \overline{f_2} \cdot f_1|_{x_i=0}
 $
@@
 swap, Rudell~\cite{RUDELL93} proposed a node overwriting strategy. A given node
-$(var(x_i), hi(x_i), lo(x_i))$ at level $i$ is overwritten by the triple
+$(var(x_i), hi(x_i), lo(x_i))$ at level $i$ is overwritten by the node triple
 \begin{center}
@@
 well as the $\text{BDD}_R$ function (for applying a relational term $R$) with a
-subsequent substitution of the place-holders with the argument variables, i.e.
+subsequent substitution of the place-holder variables with the argument variables, 
+i.e. \newline
 $\langle d_1 \leftarrow x_1, \ldots, d_n \leftarrow x_n \rangle$~\cite{BURCH90}.

### 2015-09-15 paper.tex
Revise Abstract and Introduction

@@
 This paper serves as a survey on \textit{ordered binary Decision diagrams}
-(\textit{OBDDs}). It provides an overview of the representation of OBDDs, the
-operations which can be applied efficiently, some aspects when implementing them
-as well as their limitations and how to probably overcome them by using
-alternative representations. No new results or insights are provided. \tbr
+(\textit{OBDDs}). It provides an overview of the OBDD representation, the
+operations which can be applied efficiently on OBDDs and some aspects when it
+comes to implementing OBDD packages. Moreover, their limitations and their
+application in symbolic model checking are outlined and alternative representations
+are presented to tackle these limitations. As this is just a survery, no new
+insights are provided.
 \end{abstract}
@@
 
-In computer-aided design (CAD) as well as other domains like artificial
+In computer-aided design (CAD) as well as other domains, like artificial
 intelligence or combinatorics, many problems can be formulated as Boolean
-functions. These Boolean functions then can be represented symbolically using the
-\textit{Binary Decision Diagrams} (\textit{BDDs}) introduced by Lee\cite{LEE59}
-and Akers\cite{AKERS78} in 1959 and 1978, respectively.
-
-In 1986 Bryant\cite{BRYANT86} published a paper of strong impact, describing a
-\textit{reduced} and \textit{ordered} class of BDDs, so-called \textit{ROBDDs}
-(\textit{Reduced Ordered} BDDs) but in literature mostly referred to as
+functions. These Boolean functions then can be represented symbolically using
+\textit{Binary Decision Diagrams} (\textit{BDDs}), introduced by Lee~\cite{LEE59}
+and Akers~\cite{AKERS78} in 1959 and 1978, respectively.
+
+In 1986, Bryant~\cite{BRYANT86} published a paper having strong impact, describing
+a \textit{reduced} and \textit{ordered} class of BDDs, so-called \textit{ROBDDs}
+(\textit{Reduced Ordered} BDDs), but in literature mostly referred to as
 \textit{OBDDs}. OBDDs represent Boolean functions in a \textit{canonical}
 form. This results in a compact form and very efficient tests for some properties,
-e.g. satisfiability or equivalence\cite{BRYANT86}.
+e.g. satisfiability or equivalence~\cite{BRYANT86}.
 
@@
 
-Nonetheless representing Boolean functions using OBDDs has some drawbacks. Classes
-of Boolean functions exist for which the size of the graph grows exponentially
+Representing Boolean functions using OBDDs has one drawback. Classes of
+Boolean functions exist for which the size of the graph grows exponentially
 independent of the variable ordering, e.g. integer
@@
 the problem of variable ordering~\cite{BOLLIG96}. Furthermore, alternative
-representations which e.g. relax the ordering~\cite{BRYANT95} or change the
-interpretation of nodes~\cite{BRYANT95, ANDERSEN97} were developed. Such
-representations may be beneficial for some classes of Boolean functions but mostly
-suffer from losing the canonical form and hence the desirable properties OBDDs
-are famous for.
+representations were developed which may be beneficial for some classes of Boolean
+functions but mostly suffer from losing the canonical form and, hence, the desirable
+properties OBDDs are famous for.
 
@@
 OBDD representation. The third section describes operations which can be applied
-to OBDDs efficiently. No source code is provided but those can be looked up in
-the corresponding papers easily (references are provided for each algorithm).
-Section~\ref{sec:implementational-aspects} describes aspects one has to take into
-account when implementing an OBDD package, such as what data structures to use
-or how to deal with the problem of variable ordering. Limitations of the OBDD
-representation are described in the subsequent section. The last section
-discusses alternative representations to possibly overcome the limitations of
-the OBDD representation. \tbr
+to OBDDs efficiently. Section~\ref{sec:implementational-aspects} describes aspects
+one has to take into account when implementing an OBDD package, such as what data
+structures to use or how to deal with the problem of variable ordering.
+In the subsequent sections, limitations of the OBDD representation and their usage
+in symbolic model checking are outlined. The last section discusses alternative
+representations to probably overcome the limitations mentioned before.
 
@@
 bit vectors. This representation proved successful to solve some combinatorial
-problems~\cite{BRYANT95}. \tbr \newpage
+problems~\cite{BRYANT95}. \tbr %\newpage
 
@@
 
-\clearpage
+%\clearpage
 % balance columns on last page

### 2015-09-15 paper.tex
Revise Abstract
Revise Introduction

@@
 are presented to tackle these limitations. As this is just a survery, no new
-insights are provided.
+insights are provided whatsoever.
 \end{abstract}
@@
 arguments may be omitted in some cases to keep the formulas simple, i.e. $f$
-denotes $f(x_1, \ldots, x_2)$. Nodes represent variables and have one incoming
-and two outgoing edges, each of which denotes the variable evaluates to 0 (left)
-or 1 (right), respectively. Whenever the result of $f$ is fixed, an edge points
+denotes $f(x_1, \ldots, x_n)$. Nodes represent variables and have one incoming
+and two outgoing edges, each of which denotes the evaluation of the variable to
+0 (left) or 1 (right). Whenever the result of $f$ is fixed, an edge points
 to one of the constant functions \textbf{0} or \textbf{1}. To evaluate $f$ for a
 given variable assignment, one just follows the path for the respective variable
-values. As soon as the path ends up in a constant, i.e. \textbf{0} or \textbf{1},
-this is the result of the corresponding variable assignment. Akers~\cite{AKERS78}
-already provided some techniques to simplify the BDD representation such as
-combining equivalent nodes.
+values, starting from the root. As soon as the path ends up in a constant, i.e.
+\textbf{0} or \textbf{1}, this is the result of the corresponding variable
+assignment. Akers~\cite{AKERS78} already provided some techniques to simplify the
+BDD representation, such as combining equivalent nodes.
 
@@
 $f=(A \land B) \lor (B \land C)$. To simplify the graphs, the diagram style of
-Bryant~\cite{BRYANT92} is reused, i.e. edge labels are skipped. Instead dashed
+Bryant~\cite{BRYANT92} is reused, i.e. edge labels are skipped. Instead, dashed
 and solid lines are used to represent the value of the decision variable to be 0
@@
     }
-    \caption{BDD and OBDD of $f=(A \land B) \lor (B \land C)$.}
+    \caption{BDD and OBDD of $f=(A \land B) \lor (\overline{B} \land C)$.}
     \label{fig:bdd-and-obdd}
@@
 However, Bryant~\cite{BRYANT86} introduced some restrictions concerning the
-ordering of the so-called \textit{decision variables} represented by the nodes
-of the BDD. Furthermore, Bryant\cite{BRYANT86} was able to derive a reduced
+ordering of the so-called \textit{decision variables}, represented by the nodes
+of the BDD. As a consequence, Bryant\cite{BRYANT86} was able to derive a reduced
 representation of BDDs providing the DAG in a canonical form: the
@@
 
-Figure~\ref{subfig:obdd} shows one possible ROBDD representation of
-$f$. The following section covers the details of how
-to derive the OBDD representation from a BDD. \tbr
+Figure~\ref{subfig:obdd} shows one possible ROBDD representation of $f$. The
+following section covers the details on how to derive the OBDD representation
+from a given BDD.
 

### 2015-09-15 paper.tex
Revise Operations intro and Restrict

@@
 and $m_2$, a total ordering is imposed, i.e. $var(n) < var(m_1)$ and
-$var(n) < var(m_2)$, respectively, where $var(x)$ denotes the label of node
-$x$~\cite{BRYANT86, BRYANT92}.
+$var(n) < var(m_2)$, respectively, where $var(x)$ denotes the label (variable)
+of node $x$~\cite{BRYANT86, BRYANT92}.
 
@@
 there exist classes of functions for which the OBDD complexity is exponential
-regardless of which variable ordering is chosen.
-
-Before discussing the reduction procedure, some definitions are necessary. Here
+regardless of the variable ordering chosen.
+
+Before discussing the reduction procedure, some definitions are necessary. Here,
 the original definitions of Bryant~\cite{BRYANT86, BRYANT91} are reused.
@@
 A Boolean function can be represented as an OBDD, i.e. a rooted, directed,
-acyclic graph (DAG) with a set of nodes $V$ as well as a set of edges $E$.
+acyclic graph (DAG), with a set of nodes $V$ and a set of edges $E$.
 
@@
 For each constant function, there exists only a single terminal node representing
-it, e.g. for Boolean functions, only two terminal nodes exist, representing
+it, i.e. for Boolean functions, only two terminal nodes exist, representing
 \textbf{0} and \textbf{1}, respectively. Hence, when transforming a given BDD,
@@
 holds for any other node $n$. If this is the case, $d$ can be eliminated and all
-edges are redirected to $n$.
+edges are redirected to $n$ accordingly.
 
@@
 nodes. Intuitively, this rule should be straightforward: any node having the same
-child on all outgoing edges is redundant because the actual value of the variable
+child on both outgoing edges is redundant because the actual value of the variable
 does not change the result of the function. More formally, a nonterminal node $n$
@@
 other tests, such as variable independence and functional equivalence tests, to
-be performed efficiently. \tbr
+be performed efficiently.
 
@@
 Another operation is \textit{Restrict}, restricting a given variable $x$ to a
-given (constant) value $c$, denoted $f|_{x\leftarrow c}$. Furthermore, the
+given (constant) value $c$, denoted $f|_{x=c}$. Furthermore, the
 operations \textit{Compose} and \textit{Satisfy} can be defined as graph
@@
 A variable $x_i$, argument of a function $f$, can be restricted to a given value
-$c$, denoted by $f|_{x_i\leftarrow c}$, as follows.
+$c$, denoted by $f|_{x_i=c}$, as follows.
 
@@
 OBDD depth-first, every node is checked for $var(n) = x_i$. This is done recursively
-on the given graph, generating the resultant graph This node represents
-the variable to be restricted. Now, there are two cases for $c$ to distinguish:
+on the given graph, generating the resultant graph. This node represents
+the variable to be restricted. Now, there are two cases for $c$ to be
+distinguished.
+
+\clearpage
 \begin{enumerate}
@@
 Section~\ref{sec:implementational-aspects} for details)~\cite{BRYANT92}.
-\tbr
 

### 2015-09-15 paper.tex
Revise Apply

@@
 \noindent
-Using this function, one check if $f_1$ and $f_2$ are functionally equivalent by
-computing $(f_1\overline{\oplus}f_2)\lor d$. If the result is \textbf{1}, $f_1$
-and $f_2$ are functionally equivalent.
+Using this function, one checks whether $f_1$ and $f_2$ are functionally
+equivalent by computing $(f_1\overline{\oplus}f_2)\lor d$. If the result is
+\textbf{1}, $f_1$ and $f_2$ are functionally equivalent.
 
@@
 $
-f_1 op f_2 =
-\overline{x_i} \cdot \left(f_1|_{x_i\leftarrow 0}\;op\;f_2|_{x_i\leftarrow 0}\right)
-+ x_i \cdot \left(f_1|_{x_i\leftarrow 1}\;op\;f_2|_{x_i\leftarrow 1}\right)
+f_1\;op\;f_2 =
+\overline{x_i} \cdot \left(f_1|_{x_i=0}\;op\;f_2|_{x_i=0}\right)
++ x_i \cdot \left(f_1|_{x_i=1}\;op\;f_2|_{x_i=1}\right)
 $
@@
                 If $var(n_1) = var(n_2)$, a node having $var(n_1)$ is created and
-                the procedure is applied recursively on $(lo(n_1), lo(n_2))$ and
+                the procedure is applied recursively to $(lo(n_1), lo(n_2))$ and
                 $(hi(n_1), hi(n_2))$.
@@
             \item{
-                If $var(n_1)$ is a nonterminal and $n_2$ is eiter a terminal node
+                If $var(n_1)$ is a nonterminal and $n_2$ is either a terminal node
                 or $var(n_1) < var(n_2)$, a node having $var(n_1)$ is created and
-                the procedure is applied recursively on $(lo(n_1), n_2)$ and
+                the procedure is applied recursively to $(lo(n_1), n_2)$ and
                 $(hi(n_1), n_2)$.
@@
 
-To get a reduced graph immediately, the three transformation rules are utilized
-again for each step~\cite{BRYANT92}.
+To get a reduced graph immediately, the three transformation rules are applied
+again in each step~\cite{BRYANT92}.
 
@@
 
-The full pseudocode defined by Bryant can be found in~\cite[p. 685]{BRYANT86}. \tbr
+The full pseudocode defined by Bryant can be found in~\cite[p. 685]{BRYANT86}.
 

### 2015-09-15 paper.tex
Revise Compose

@@
 a function $d(x_1, \ldots, x_n)$ for \textit{Don't cares}:
-
+\begin{center}
 $d(x_1, \ldots, x_n)=
@@
 $
+\end{center}
 
@@
 $O\left(|V_{f_1}|^2 \cdot |V_{f_2}|^2\right)$. Therefore, Bryant~\cite{BRYANT86}
-already proposed to extend the \textit{Apply} operation to support ternary
-operations and then use the equivalent \textit{if-then-else} ($ITE$)
-representation:
+proposed to extend the \textit{Apply} operation to support ternary operations and
+then use the equivalent \textit{if-then-else} ($ITE$) representation:
 \begin{center}
@@
 This formulation is based upon the fact, that every node $n$ of an OBDD represents
-a Boolean function and is denoted by a triple $(var(n), hi(n), lo(n))$. Brace et
-al.~\cite{BRACE90} called the variable associated with $n$, i.e. $x_n$ the
-\textit{top variable} of $n$.
+a Boolean function and is denoted by a triple $(var(n), hi(n), lo(n))$.
+%Brace et al.~\cite{BRACE90} called the variable associated with $n$ the
+%\textit{top variable} of $n$.
 
@@
 two function graphs. But this approach only applicable under restricted conditions.
-\tbr
 

### 2015-09-15 paper.tex
Revise Satisfy

@@
 \textit{Satisfy-one} traverses the OBDD depth-first and uses backtracking until
-a terminal node of the constant function \textbf{1} is encountered. If such a
-node is found, the corresponding Boolean function is satisfiable and true is
-returned. Furthermore, an array representing the satisfying variable assignment
+the terminal node of the constant function \textbf{1} is encountered. If this node
+is reached, the corresponding Boolean function is satisfiable and \textbf{True}
+is returned. Furthermore, an array representing the satisfying variable assignment
 is assigned. The time complexity for a reduced graph is linear, i.e.
@@
 
-\textit{Satisfy-all} also traverses the OBDD depth-first. Every time the terminal
-node having value \textbf{1} is traversed, it is printed. The recursive procedure
-terminates as soon as the terminal node having value \textbf{0} is reached. This
-results in a time complexity of $O(|V_f| \cdot |S_f|)$, whereas $S_f$ denotes the
-satisfying set~\cite{BRYANT86}. \tbr
+\textit{Satisfy-all} also traverses the OBDD depth-first. Each time the terminal
+node having value \textbf{1} is traversed, the element corresponding to the current
+path is printed. The recursive procedure terminates as soon as the terminal node
+having value \textbf{0} is reached. This results in a time complexity of
+$O(|V_f| \cdot |S_f|)$, whereas $S_f$ denotes the satisfying set~\cite{BRYANT86}.
 

### 2015-09-15 paper.tex
Revise Implementational aspects
Revise Efficient data structures

@@
 implementing an OBDD framework efficiently, i.e. efficient data structures, the
-$ITE$ operator, how to possibly approach the variable ordering problem and
-\textit{garbage collection}. \tbr
+$ITE$ operator, \textit{garbage collection} and one possible approach to the variable
+ordering problem.
 
@@
 value for a given key in (amortized) constant time. Collisions, i.e. equivalent
-value for different keys, are resolved by chaining all associated values in a 
+values for different keys, are resolved by chaining all associated values in a 
 linked list~\cite{BRACE90}.
@@
 of nodes. If so, the associated value $m$ is used instead of computing the
-subgraph~\cite{BRYANT86, BRYANT92}.
+subgraph again~\cite{BRYANT86, BRYANT92}.
 
@@
 operation~\cite{BRYANT92}. This is also referred to as
-\textit{unique-table}~\cite{BRACE90} imposes the canonical form of an OBDD. If a
+\textit{unique-table}~\cite{BRACE90} and imposes the canonical form of an OBDD. If a
 new node is generated, a corresponding entry is created in the unique-table.
 
-There exists a special version of hash tables, so-called
-\textit{hash-based caches}, which do not resolve collisions, i.e. overwrite
-existing entries, and issue a cache miss if a looked up element is not found.
-As will be seen, \textit{Hash-based caches} are useful to provide an efficient
-implementation of the $ITE$ operator. \tbr
+There exists a special type of hash tables, so-called
+\textit{hash-based caches}, which do not resolve collisions, i.e. they overwrite
+existing entries and issue a \textit{cache miss} if a looked up element is not
+found. As will be seen, \textit{hash-based caches} are useful to provide an
+efficient implementation of the $ITE$ operator. \tbr
 
@@
 present, a complement edge can be used to connect $f$ and thus avoid
-constructing the intermediate node $\overline{f}$. Also one of the two terminal
-nodes (e.g. \textbf{1}) can be removed by just using a complement edge to connect
-the other one (e.g. \textbf{0})~\cite{BRACE90}.
+constructing the intermediate node $\overline{f}$. Using this, one of the two
+terminal nodes (e.g. \textbf{1}) can be removed too, by using a complement edge
+to connect the other one (e.g. \textbf{0})~\cite{BRACE90}.
 
@@
 respectively. Moreover, let $in(x)$ be the (one) incoming edge of node $x$ and
-$\overline{c}$ the complement of $c$. Then, the following table sums up the
-four pairs identified by Brace et al.~\cite{BRACE90}.
+$\overline{c}$ be the complement of $c$. In the following, these four pairs,
+identified by Brace et al.~\cite{BRACE90}, are summarized:
 
@@
 
-Also, the 1-edge of each node must always be a regular one to guarantee a
+Also, the 1-edge of each node must always be a regular one to guarantee the
 canonical form.
@@
 complement functions in constant time (because $f$ and $\overline{f}$ use the same
-node). According to Brace et al.~\cite{BRACE90} this reduces the total runtime
-to form the OBBDD by almost a factor of 2 (for 12 examples).
+node). According to Brace et al.~\cite{BRACE90}, this technique reduces the total
+runtime to form the OBDD by almost a factor of 2 (for 12 examples).
 
@@
 i.e. linking the collision chain of each node to the corresponding unique-table
-entry, to improve memory usage. \tbr
+entry, in order to improve memory usage.
 

### 2015-09-15 paper.tex
Revise ITE operator

@@
 \textit{caching computed-table}, to implement it, exploiting a high locality of
-reference and consuming less memory.
-
-Compared to using a normal hash table, this results in more recursive calls to 
-$ITE$ (because of cache misses) but less time spent in garbage collection
-(described in Section~\ref{subsec:garbage-collection}) and look-ups (because no
-collision chains need to be searched).
+reference and thus consuming less memory.
+
+Compared to using a normal hash table, this results in more recursive calls 
+(because of cache misses) but less time spent in garbage collection (described
+in Section~\ref{subsec:garbage-collection}) and look-ups (because no collision
+chains need to be searched).
 
@@
 logical implication tests more efficiently by avoiding the construction of
-intermediate nodes. \tbr
+intermediate nodes.
 

### 2015-09-15 paper.tex
Revise ITE operator
Revise Garbage collection

@@
 This formulation is based upon the fact, that every node $n$ of an OBDD represents
-a Boolean function and is denoted by a triple $(var(n), hi(n), lo(n))$.
+a Boolean function and is denoted by a triple $(var(n), hi(n), lo(n))$. $var(x)$
+here denotes the top-most variable among the three OBDDs $x$, $h$ and $l$.
 %Brace et al.~\cite{BRACE90} called the variable associated with $n$ the
@@
 introduced by Brace, Bryant and Rudell~\cite{BRACE90}. The garbage collection was
-based upon the reference counting technique, i.e. every node of the OBDD had a
+based upon the reference counting technique, i.e. every node of the OBDD has its own
 reference count. This reference count represents two kinds of references, (1)
@@
 
-Today, garbage collecion is a well-known topic. It is used to free up unnecessary
-allocated or unused memory. To accomplish this, the garbage collection has to know
-about which parts of the memory are \textit{dead} and thus can be freed. A node
+Today, garbage collecion is a well-known topic. It is used to free memory of
+objects that are no longer in use. To accomplish this, the garbage collection has
+to know which parts of the memory are \textit{dead} and thus can be freed. A node
 of an OBDD is called \textit{dead} if the reference count of it is equal to zero.
@@
 The garbage collection is executed whenever the load factor (the ratio between
-number of keys stored and number of available bins) exceeds 4 after an insertion
+number of stored keys and number of available bins) exceeds 4 after an insertion,
 and $10\%$ of all OBDD nodes are dead. A garbage collection run deletes all entries
@@
 memory overflow of computed- and unique-table, although, the $10\%$-threshold of
-dead nodes is used again. In any other case, the garbage collection gives up (and,
+dead nodes is used here again. In any other case, the garbage collection gives up (and,
 of course, frees all intermediate nodes)~\cite{BRACE90}.
@@
 As stated by Brace et al.~\cite{BRACE90}, this provides effective memory usage
-at very low run-time costs, i.e. $3-7\%$ of the total runtime. \tbr
+at very low runtime costs, i.e. $3-7\%$ of the total runtime.
 

### 2015-09-15 paper.tex
Revise Dynamic variable ordering

@@
 For an OBDD package it is desirable to provide the possibility to automatically 
-choose a good, though not optimal variable order for arbitrary functions. Of
+choose a good, though not optimal, variable ordering for arbitrary functions. Of
 course, there exist heuristics to order variables before processing but
 Rudell~\cite{RUDELL93} proposed to integrate the determination and maintainance
-of the variable order into the OBDD package itself. This approach to the variable
+of the variable ordering into the OBDD package itself. This approach to the variable
 ordering problem is called \textit{dynamic variable ordering} and increases the
-robustness of an OBDD package by removing the necessity of choosing a proper
-variable ordering beforehand and thus running the risk of spacing out.
+robustness of an OBDD package by removing the necessity to choose a proper 
+variable ordering beforehand and thus, running the risk of spacing out.
 
@@
 
-Periodically means, that a reordering is not performed by user requests but at
-appropriate points, determined by the OBDD package itself and so eliminate any
-user responsibility.
+Periodically means, that a reordering is not performed on user requests but at
+appropriate points of time, determined by the OBDD package itself and so
+eliminate any user responsibility.
 
@@
 is located on the layer below in the OBDD. This is done until $x$ is the next to
-last variable in the OBDD. Then, this is done in the other way round, i.e. swapping
+last variable in the OBDD. Then, this is done the other way round, i.e. swapping
 $x$ with its predecessor variable until $x$ becomes the top variable of the OBDD.
@@
 is marked red, i.e. the two swapped variables are marked. The variable pairs
-swapped during the second phase, i.e. moving the $x$ to its ideal position (4),
-are marked green.
+swapped during the second phase, when moving $x$ to its ideal position (4), are
+marked green.
 
@@
 swapping the processed variable. As a result, the variable is always swapped
-towards the closer variable position first.
+towards the closer end (top or bottom) first.
 
@@
 utilizing the approach described above. The sifting algorithm has one big advantage
-over other similar algorithms, like \textit{window permutation algorithm}, namely
+over other similar algorithms, like \textit{window permutation algorithm}, namely,
 variables can move many levels because all positions are evaluated, regardless of
@@
 According to Rudell~\cite{RUDELL93}, $O(n^2)$ swaps of adjacent levels are
-necessary. Moreover, each swap has a complexity proportional to the with of the
+necessary. Moreover, each swap has a complexity proportional to the width of the
 OBDD. Thus, Rudell~\cite{RUDELL93} advises to terminate whenever the OBDD doubles
-its size. On the one hand, in contrast to heuristics and similar approaches, i.e.
+its size. On the one hand, in contrast to heuristics and similar approaches, e.g.
 the window permutation algorithm, the dynamic variable ordering using the sifting
@@
 ordering using the sifting algorithm was able to complete operations which were
-not completed without it. Furthermore, the resulting OBDDs were significantly
-smaller~\cite{RUDELL93}.
+not completed using different approaches. Furthermore, the resulting OBDDs were
+significantly smaller~\cite{RUDELL93}.
 
@@
 \textit{negate-input-} and \textit{negate-else-edges} cannot be used without
-loosing the (desired) local complexity of variable swapping.
+loosing the (desirable) local complexity of variable swapping.
 
@@
 For example, Figure~\ref{fig:min-obdd} shows the minimized, and thus optimal, OBDD
-representation of \newline $f=\left(A \land B\right) \lor \left(B \land C\right)$
-which, compared to the OBDD shown in Figure~\ref{subfig:obdd}, consists of less
-nodes.
+representation of \newline
+$f=\left(A \land B\right) \lor \left(\neg B \land C\right)$ which, compared to
+the OBDD shown in Figure~\ref{subfig:obdd}, consists of less nodes.
 
@@
 \end{figure}
-\tbr
 

### 2015-09-15 paper.tex
Revise Introduction again

@@
 Figure~\ref{subfig:bdd} shows an example of a BDD representing the Boolean function
-$f=(A \land B) \lor (B \land C)$. To simplify the graphs, the diagram style of
+$f=(A \land B) \lor (\neg B \land C)$. To simplify the graphs, the diagram style of
 Bryant~\cite{BRYANT92} is reused, i.e. edge labels are skipped. Instead, dashed
@@
     }
-    \caption{BDD and OBDD of $f=(A \land B) \lor (\overline{B} \land C)$.}
+    \caption{BDD and OBDD of $f=(A \land B) \lor (\neg B \land C)$.}
     \label{fig:bdd-and-obdd}
@@
 ordering $\pi$. Bryant~\cite{BRYANT91} mainly focused on the set of integer
-multipliers having word size $n$.
+multipliers (having word size $n$).
 

### 2015-09-15 paper.tex
Revise Limitations

@@
 
-Nevertheless, researches put some effort in improving this lower bound. In 2001,
-Woelfel~\cite{WOELFEL01} published a new bound lower bound as well as a first 
-nontrivial upper bound for the size of OBDDs representing $n$-bit integer
-multiplication by utilizing the concept of \textit{universal hashing}. Therefore,
-he utilized a universal family of hash functions to improve the exponential lower
-bound on the OBDD-size of integer multiplication to
-$2^{\lfloor\frac{n}{2}\rfloor}/96$. Moreover, Woelfel~\cite{WOELFEL01} proved
-the upper bound for the OBDD-size of computing the middle bit of an $n$-bit
-multiplication to be $7/3 \cdot 2^{\frac{4 \cdot n}{3}}$.
-
-Nevertheless, all lower bounds for $n$-bit integer multiplication are exponential
+Nevertheless, researchers put some effort in improving this lower bound. In 2001,
+Woelfel~\cite{WOELFEL01} published a new lower bound as well as a first nontrivial
+upper bound for the size of OBDDs representing $n$-bit integer multiplication by
+using the concept of \textit{universal hashing}. Therefore, he utilized a
+universal family of hash functions to improve the exponential lower bound on the
+OBDD-size of integer multiplication to $2^{\lfloor n/2 \rfloor}/96$.
+Moreover, Woelfel~\cite{WOELFEL01} proved the upper bound for the OBDD-size of
+computing the middle bit of an $n$-bit multiplication to be
+$7/3 \cdot 2^{\left(4 \cdot n \right) / 3}$.
+
+However, both lower bounds for $n$-bit integer multiplication are exponential
 in the number of nodes. Especially because of the high practical relevance of
-$n$-bit integer multiplication this is one of the main drawbacks of the OBDD
+$n$-bit integer multiplication, this is one of the main drawbacks of the OBDD
 representation~\cite{BRYANT86, BRYANT91, WOELFEL01}.
 
-Another topic researchers put much effort into was the variable ordering problem.
-Section~\ref{subsec:dynamic-variable-ordering} already described one possible
+Another topic tackled by some research teams was the variable ordering problem.
+Section~\ref{subsec:dynamic-variable-ordering} already outlined one possible
 approach to this problem but there are many other approaches such as heuristics
@@
 
-Bollig and Wegener~\cite{BOLLIG96} studied the complexity of improving the
-variable ordering. They investigated this problem by analyzing the following
-problem: Given an OBDD representing $f$ and a size bound $s$. Does an OBDD exist,
-also representing $f$ and respecting an arbitrary variable ordering, having at
-most $s$ nodes? "Unfortunately", they were able to prove this problem to be 
-NP-complete. Hence, improving the variable ordering of OBDDs is
-NP-complete~\cite{BOLLIG96}. \tbr
+Consequently, Bollig and Wegener~\cite{BOLLIG96} studied the complexity of
+improving the variable ordering. They investigated this problem by analyzing the
+following problem: Given an OBDD representing $f$ and a size bound $s$. Does an
+OBDD exist, also representing $f$ and respecting an arbitrary variable ordering,
+having at most $s$ nodes? "Unfortunately", they were able to prove this problem
+to be NP-complete. Hence, improving the variable ordering of OBDDs is
+NP-complete too~\cite{BOLLIG96}.
 

### 2015-09-15 paper.tex
Revise Symbolic model checking

@@
 \textit{symbolic model checking}, introduced by Burch et al.~\cite{BURCH90} in
-1990. \textit{Symbolic}, because the state space is not represented explicitly
-but symbolically by utilizing OBDDs to represent $\mu$-Calculus formulas.
-In~\cite{BURCH90}, Burch, Clarke, McMillan, Dill and Hwang derived an efficient
+1990. Symbolic, because the state space is not represented explicitly but
+symbolically by utilizing OBDDs to represent $\mu$-Calculus formulas.
+In~\cite{BURCH90}, Burch, Clarke, McMillan, Dill and Hwang derived efficient
 decision procedures for CTL model checking as well as satisfiability of LTL
@@
 state space symbolically. Although OBDDs cannot avoid a state explosion in general,
-many practically relevant systemes with large state spaces can be
+many practically relevant systems with large state spaces can be
 verified~\cite{BURCH90, MCMILLAN92}. Experiments~\cite{BURCH90} showed that
@@
 
-Burch et al.~\cite{BURCH90} and McMillan~\cite{MCMILLAN92} extend Bryant's OBDD
+Burch et al.~\cite{BURCH90} and McMillan~\cite{MCMILLAN92} extended Bryant's OBDD
 representation~\cite{BRYANT86} by quantification over Boolean variables (QBF) and
@@
 substitution of $x$ by $y$ (denoted $f\langle x\leftarrow y\rangle$), is computed
-using the following equivalence:
+using the following definition:
 \begin{center}
 $
-f\langle x\leftarrow y\rangle \equiv
+f\langle x\leftarrow y\rangle :=
 \exists x\lbrack\left(x \Leftrightarrow y\right) \land f\rbrack
@@
 
-Before discussing the symbolic model checking algorithms $\text{BDD}_f$ and
-$\text{BDD}_R$, the used $\mu$-Calculus has to be described. For a given finite
-signature $\mathcal{S}$, each symbol is either an individual or a predicate
-variable having positive arity. Burch et al.~\cite{BURCH90} considered two
-syntactic categories, i.e. \textit{formulas} and \textit{relational terms}, having
-the following forms.
-
-\noindent
+Before discussing the symbolic model checking algorithm, the used $\mu$-Calculus
+has to be defined. For a given finite signature $\mathcal{S}$, each symbol is
+either an individual or a predicate variable having positive arity. Burch et
+al.~\cite{BURCH90} considered two syntactic categories, i.e. \textit{formulas}
+and \textit{relational terms}, having the following forms.
+
+\vskip5pt\noindent
 \textbf{Formulas:}~\cite{BURCH90}
@@
     \item{
-        $\neg f$, $f \lor g$, $\exists z\left( f \right)$; $z\in\mathcal{S}$ is
-        an individual variable, $f$ and $g$ are formulas.
+        $\neg f$, $f_1 \lor f_2$, $\exists z\left( f \right)$; $z\in\mathcal{S}$ is
+        an individual variable, $f_1$ and $f_2$ are formulas.
     }
@@
 
-Finally, after defining the $\mu$-Calculus formally, the model checking algorithms
-can be described. They operate on the limited Boolean domain, i.e.
-$D = \lbrace 0, 1 \rbrace$ but Burch et al.~\cite{BURCH90} also described how any
-\textit{finite} domain can be encoded into this limited Boolean domain. This last
-fact generalizes the introduced algorithms to any \textit{finite} domain.
-
-They described two function representing the algorithm:
+Finally, after defining the $\mu$-Calculus formally, the model checking algorithm
+can be described, which operates on the limited Boolean domain, i.e.
+$D = \lbrace 0, 1 \rbrace$. However, Burch et al.~\cite{BURCH90} also described
+how any \textit{finite} domain can be encoded into this limited Boolean domain.
+This last fact generalizes the introduced algorithm to any \textit{finite} domain.
+
+They described two functions representing the algorithm:
 $\text{BDD}_f \left( f, I_P \right)$ and $\text{BDD}_R \left( R, I_P \right)$,
@@
 $d_1, \ldots, d_n$ are used to represent the values of the relational variables
-$\in I_P$. $d_i$ represents the $i^{th}$ argument of a relation and an $n$-ary
+in $I_P$. $d_i$ represents the $i^{th}$ argument of a relation and an $n$-ary
 relation holds for arguments $y_1, \ldots, y_n$ iff the interpretation
@@
 derived from the OBDD semantics and the $\mu$-Calculus, utilizing the graph-based
-algorithms of Bryant~\cite{BRYANT86} (also see Section~\ref{sec:operations}) as
+algorithms of Bryant~\cite{BRYANT86} (see Section~\ref{sec:operations}) as
 well as the $\text{BDD}_R$ function (for applying a relational term $R$) with a
@@
 
-$\text{\textbf{EX}} f$ is satisfied for a state $s$ iff a state satisfying $f$
-exists and $R \left( s, t \right)$ holds. In terms of $\mu$-Calculus relational
-terms, $\text{\textbf{EX}} f$ can be formulated as
+$\text{\textbf{EX}} f$ is satisfied for a state $s$ iff a state $t$ satisfying
+$f$ exists and $R \left( s, t \right)$ holds. In terms of $\mu$-Calculus
+relational terms, $\text{\textbf{EX}} f$ can be formulated as
 $\lambda u \lbrack \exists v \left( f(v) \land R \left( u, v \right) \right) \rbrack$~\cite{BURCH90}.
@@
 
-Utilizing these equivalent representations, one efficient CTL model checking.
-Moreover, Burch, Clarke, McMillan and Dill~\cite{BURCH90-2} described how to use
-these methods to do \textit{fair} CTL model checking. \tbr
+Utilizing these equivalent representations, efficient CTL model checking can be
+done. Moreover, Burch, Clarke, McMillan and Dill~\cite{BURCH90-2} described how
+to use these methods to do \textit{fair} CTL model checking.
 

### 2015-09-15 paper.tex
Revise Alternative representations intro

@@
 In this last section, some alternative representations of BDDs are discussed
-briefly, e.g. BMDs or BEDs. These representations were introduced primarily
+briefly, i.e. FDDs, ZBDDs and BEDs. These representations were introduced primarily
 because the OBDD representation is insufficient for some problems of practical
@@
 \textit{Binary Expression Diagrams} introduced by Andersen and
-Hulgaard~\cite{ANDERSEN97}. \tbr
+Hulgaard~\cite{ANDERSEN97} in 1997.
 

### 2015-09-15 paper.tex
Revise FBDDs

@@
 In this last section, some alternative representations of BDDs are discussed
-briefly, i.e. FDDs, ZBDDs and BEDs. These representations were introduced primarily
+briefly, i.e. FBDDs, ZBDDs and BEDs. These representations were introduced primarily
 because the OBDD representation is insufficient for some problems of practical
@@
 ordering chosen, e.g. integer multiplication (see Section~\ref{sec:limitations}).
-One way researchers tried to overcome this issue is relaxing the variable ordering
-requirement. \textit{Free BDDs} (\textit{FDDs}) is one proposed approach. FDDs
-do not require the variables in the BDD to be ordered but, instead, require that
-each variable only occurs once along any path (starting from the root node).
-Hence, FDDS are also referred to as \textit{1-time branching programs}. Moreover,
-FDDs have the same desirable algorithmic properties as OBDDs~\cite{BRYANT95}. \tbr
+One way researchers tried to overcome this issue is by relaxing the variable
+ordering requirement. \textit{Free BDDs} (\textit{FBDDs}) is one proposed
+approach. FBDDs do not require the variables in the BDD to be ordered but,
+still, require that each variable only occurs once along any path (starting from
+the root node). Hence, FBDDS are also referred to as
+\textit{1-time branching programs}. Moreover, FBDDs have the same desirable
+algorithmic properties as OBDDs~\cite{BRYANT95}.
 

### 2015-09-15 paper.tex
Revise ZBDDs

@@
 Boolean function over $n$ variables. If the corresponding bit vector is in the
-set, the \textit{Zero-Suppressed BDDs} (\textit{ZBDDs}) yields 1 for a given
+set, the \textit{Zero-Suppressed BDD} (\textit{ZBDD}) yields 1 for a given
 variable assignment.
 
-The bit vector are \textit{sparse} in the sense that the set contains much less
-than $2^n$ vectors and the bit vectors themselves have many zero elements. ZBDDs
-are similar to OBDDs but they use a different reduction rule, exploiting the
-sparse bit vectors. In ZBDDs, nodes can be eliminated whenever setting the
-corresponding variable to 1 yields 0 as result, which is often the case for sparse
-bit vectors. This representation proved successful to solve some combinatorial
-problems~\cite{BRYANT95}. \tbr %\newpage
+Bit vectors are \textit{sparse} in the sense that the set of bit vectors contains
+much less than all possible $2^n$ bit vectors and the bit vectors themselves
+contain many zero elements. ZBDDs are similar to OBDDs but they use a different
+reduction rule, exploiting the sparseness of the bit vectors. In ZBDDs, nodes can be
+eliminated whenever setting the corresponding variable to 1 yields 0 as result,
+which is often the case for sparse bit vectors. This representation proved
+successful to solve some combinatorial problems~\cite{BRYANT95}.
 
@@
 
-Analogous to OBDDs and FDDs, OBEDs and FBEDs can be defined, whereas OBEDs are
+Analogous to OBDDs and FBDDs, OBEDs and FBEDs can be defined, whereas OBEDs are
 OBDDs without operator nodes. The reduction rule of BEDs is similar to the one

### 2015-09-15 paper.tex
Revise BEDs

@@
 contain many zero elements. ZBDDs are similar to OBDDs but they use a different
-reduction rule, exploiting the sparseness of the bit vectors. In ZBDDs, nodes can be
-eliminated whenever setting the corresponding variable to 1 yields 0 as result,
+reduction rule, exploiting the sparseness of the bit vectors. In ZBDDs, nodes can
+be eliminated whenever setting the corresponding variable to 1 yields 0 as result,
 which is often the case for sparse bit vectors. This representation proved
@@
 
-In 1997, Andersen and Hulgaard~\cite{ANDERSEN97} proposeded extending OBDDs with
+In 1997, Andersen and Hulgaard~\cite{ANDERSEN97} proposed to extend OBDDs with
 a third type of nodes representing \textit{operators}. The resulting
@@
 expressions are terminal nodes. The BED may be simplified afterwards. These steps
-can be repeated to reduce the BED over and over again, and simplifying some tests,
-e.g. tautology testing. It is also possible to move all variable up in a single
-step. Nevertheless, reducing the BED step-by-step can exploit structural
+can be repeated to reduce the BED over and over again, and thus to simplify some
+tests, e.g. tautology testing. It is also possible to move all variables up in a
+single step. Nevertheless, reducing the BED step-by-step can exploit structural
 information~\cite{ANDERSEN97}.
 
-In contrast to OBDDs, BEDs are \textit{exponentially more succinct} because
-arbitrary circuits can be translated into BEDs. As a consequence, the BED
-representation suffices to solve integer multiplication using a quadratic number
-of nodes~\cite{ANDERSEN97}.
+In contrast to OBDDs, BEDs are \textit{exponentially more succinct}, and arbitrary
+circuits can be translated into BEDs. As a consequence, the BED representation
+suffices to solve integer multiplication using a quadratic number of
+nodes~\cite{ANDERSEN97}.
 
 Analogous to OBDDs and FBDDs, OBEDs and FBEDs can be defined, whereas OBEDs are
-OBDDs without operator nodes. The reduction rule of BEDs is similar to the one
+OBDDs without operator nodes. The reduction rules of BEDs are similar to the rules
 of OBDDs but extended with an additional rule for operator nodes. For any operator
-nodes $n$ it has to hold that $lo(n)$ and $hi(n)$ are nonterminals in order to
+node $n$ it has to hold that $lo(n)$ and $hi(n)$ are nonterminals in order to
 guarantee reducedness. Apart from that, before generating a new node, an existence
-check is performed. If there exists an equivalent node, this node is reused.
-Otherwise the new node is created and considered in the next check. However, even
-if a BED is reduced, canonicity is not guaranteed (in contrast to
-OBDDs)~\cite{ANDERSEN97}.
+check is performed. If there exists an equivalent node, this node is recycled.
+Otherwise, the new node is created and considered in the next check. However, in
+contrast to OBDDs, reducedness of BEDs does not imply
+canonicity~\cite{ANDERSEN97}.
 
 Andersen and Hulgaard~\cite{ANDERSEN97} also introduced some checks especially
-for operator nodes in order to reduce the BED-size even further. Moreover a
+for operator nodes in order to reduce the BED-size even further. Moreover, a
 detailed description of how to convert a BED into an equivalent OBDD is provided,
@@
 The \textit{up\_one} operation can be implemented with expected runtime of $O(n)$,
-where $n$ denotes the number reachable of nodes from a given node. Nonetheless,
+where $n$ denotes the number of reachable nodes from a given node. Nonetheless,
 because BEDs are exponentially more succinct than OBDDs, the complete
@@
 Andersen and Hulgaard~\cite{ANDERSEN97} concluded that \textit{up\_all} is a
-memory efficient approach to construct OBDDs in contrast to the OBDD construction
-from scratch, without actually increasing the runtime. \textit{up\_one} is used
-to implement existential quantification and substitution efficiently, having
-linear runtime complexity in the size of the BED. Reconsidering symbolic model
-checking, BEDs can be used to evaluate fixed point computations directly without
-converting it to an OBDD. Moreover, BEDs can perform fixed point computations more
-efficiently because of their succinct representation~\cite{ANDERSEN97}. \tbr
+memory efficient alternative to construct OBDDs compared to the OBDD construction
+from scratch using \textit{Apply}, without actually increasing the runtime.
+\textit{up\_one} is used to implement existential quantification and substitution
+efficiently, having linear runtime complexity in the size of the BED.
+Reconsidering symbolic model checking, BEDs can be used to evaluate fixed point
+computations directly without converting to OBDDs. Moreover, BEDs can perform
+fixed point computations more efficiently because of their succinct
+representation~\cite{ANDERSEN97}.
 

### 2015-09-15 paper.tex
Revise Abstract

@@
 \begin{abstract}
-This paper serves as a survey on \textit{ordered binary Decision diagrams}
+This paper serves as survey on \textit{ordered binary Decision diagrams}
 (\textit{OBDDs}). It provides an overview of the OBDD representation, the
@@
 application in symbolic model checking are outlined and alternative representations
-are presented to tackle these limitations. As this is just a survery, no new
+are presented to tackle these limitations. As this is only a survey, no new
 insights are provided whatsoever.
